---
title: "Dep Map analysis stats generator"
output:
  pdf_document:
    toc: TRUE
params:
  release: "19Q3"
---
This document loads the data generated in depmap_generate, and then generates and stores stats for other RMDs to load

##Load libraries
```{r load_block, echo=TRUE, message=FALSE, warning=FALSE}
#not sure which exactly of these I need
library(tidyverse)
library(lubridate)
library(here)
#library(readxl)
#library(janitor)
library(feather)
library(corrr)
library(purrr)
#library(viridis)
#library(enrichR)
#library(glue)
library(moderndive)
#library(rmarkdown)
#library(gridExtra)
#library(knitr)
#library(pander)
#library(vroom)
library(beepr) #long analysis; get some coffee, and comeback when ready

#clear environment
#rm(list=ls()) 

#how long?
start_time <- Sys.time()
```

##import
```{r import}
##BROAD
achilles <- read_feather(here::here("data", paste0(params$release, "_achilles.feather")))
achilles_long <- achilles %>% gather("gene", "dep_score", -X1)
achilles_cor <- read_feather(here::here("data", paste0(params$release, "_achilles_cor.feather"))) 
class(achilles_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df

achilles_cor_long <- achilles_cor %>% 
  stretch() #observations across 3 variables (x, y, r); longest step, but cannot get cor_long.feather to load properly

#EXPRESSION(BROAD), need this to get cell line names
expression_id <- read_feather(here::here("data", paste0(params$release, "_expression_id.feather"))) 

expression_join <- expression_id %>% 
  rename(X1 = dep_map_id) %>% 
  select(X1, stripped_cell_line_name, lineage)

gene_summary <- read_feather(path = here::here("data", "gene_summary.feather"))
```

#Permutation tests
A permutation test involves permuting one or more variables in a data set before performing the test, in order to break any existing relationships and simulate the null hypothesis. In this case, we will break the relationshipe between gene-gene pairs and the correlation values. We can then compare the true statistic (mean correlation) to the generated distribution of null statistics (fake means). This will give a better idea of where to draw a threshold of a "signficiant correlation" for these analyses. Recalculate here for statistical threshold cutoffs.

```{r permutation}
  virtual_achilles <- achilles_cor_long %>% #achilles_cor_long already has all of the variables in a long format
  filter(!is.na(r)) %>%   
  rep_sample_n(size = 20000, reps = 1000) %>% #larger sample size, less error (but only 625 sets, and we're mimicking 1000?, but 310M combinations, so probably OK)
  group_by(replicate) %>% 
  summarize(mean = mean(r), max = max(r), min = min(r), sd = sd(r)) #how to handle + vs. - correlation?

#The mean may not be the best measure of centrality due to its sensitivity to outliers. The example data used here seems asymmetric and in that case the median is a better measure of centrality.
#https://github.com/JoachimGoedhart/PlotsOfDifferences

ggplot(virtual_achilles, aes(x = sd)) +
  geom_histogram(color = "white") +
  labs(x = "SD r^2 value from virtual achilles", 
       title = "Distribution of mean r^2 values") 

mean_virtual_achilles <- mean(virtual_achilles$mean)
sd_virtual_achilles <- mean(virtual_achilles$sd)

sd_threshold <- 3

achilles_upper <- mean_virtual_achilles + sd_threshold*sd_virtual_achilles
achilles_lower <- mean_virtual_achilles - sd_threshold*sd_virtual_achilles
```

#save
```{r}
saveRDS(sd_threshold, file = here::here("data", "sd_threshold.rds"))
saveRDS(achilles_lower, file = here::here("data", "achilles_lower.rds"))
saveRDS(achilles_upper, file = here::here("data", "achilles_upper.rds"))
saveRDS(mean_virtual_achilles, file = here::here("data", "mean_virtual_achilles.rds"))
saveRDS(sd_virtual_achilles, file = here::here("data", "sd_virtual_achilles.rds"))
```

#remove and reload
```{r}
rm(mean_virtual_achilles, sd_virtual_achilles, sd_threshold, achilles_lower, achilles_upper)

sd_threshold <- readRDS(file = here::here("data", "sd_threshold.rds"))
achilles_lower <- readRDS(file = here::here("data", "achilles_lower.rds"))
achilles_upper <- readRDS(file = here::here("data", "achilles_upper.rds"))
mean_virtual_achilles <- readRDS(file = here::here("data", "mean_virtual_achilles.rds"))
sd_virtual_achilles <- readRDS(file = here::here("data", "sd_virtual_achilles.rds"))
```

