---
title: "Dep Map correlation analysis report generator"
output:
  pdf_document:
    toc: TRUE
params:
  release: "19Q3"
---
This is the report generator code.

##Load libraries
```{r load_block, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(here)
library(readxl)
library(janitor)
library(feather)
library(corrr)
library(purrr)
library(viridis)
library(enrichR)
library(glue)
library(moderndive)
library(rmarkdown)
library(gridExtra)
library(knitr)
library(pander)
library(vroom)
library(beepr) #long analysis; get some coffee, and comeback when ready

#clear environment
#rm(list=ls()) 

#how long?
start_time <- Sys.time()
```

#define functions
```{r}
#pathway enrichment analysis loop function
enrichr_loop <- function(gene_list, databases){
  if(is_empty(gene_list)){
    flat_complete <- NULL
    return(flat_complete)
  } else {
    flat_complete <- as_tibble()
    for(lib in databases){
      enriched <- enrichr(gene_list, lib)
      
      flat <- flatten_dfc(enriched) %>% 
        mutate(enrichr = lib)
      
      flat_complete <- flat_complete %>% 
        bind_rows(flat)
    }
    flat_complete <- flat_complete %>% 
      arrange(Adjusted.P.value) %>% 
      select(enrichr, Term, Overlap) 
    
    flat_complete$enrichr <- str_replace_all(flat_complete$enrichr, "\\_", " ")
    flat_complete$Term <- str_replace_all(flat_complete$Term, "\\_", " ")
    return(flat_complete)
  }
}
```

##import
Import Broad Insititute and Sanger Depedencay data; added Broad cell line expression data (Aug 21, 2019).
```{r import}
##BROAD
achilles <- read_feather(here::here("data", paste0(params$release, "_achilles.feather")))
achilles_long <- achilles %>% gather("gene", "dep_score", -X1)
achilles_cor <- read_feather(here::here("data", paste0(params$release, "_achilles_cor.feather"))) 
class(achilles_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df

achilles_cor_long <- achilles_cor %>% 
  stretch() #observations across 3 variables (x, y, r); longest step, but cannot get cor_long.feather to load properly

#EXPRESSION(BROAD), need this to get cell line names
expression_id <- read_feather(here::here("data", paste0(params$release, "_expression_id.feather"))) 

expression_join <- expression_id %>% 
  rename(X1 = dep_map_id) %>% 
  select(X1, stripped_cell_line_name, lineage)

gene_summary <- read_feather(path = here::here("data", "gene_summary.feather"))
```

#stats; import saved files generated in depmap_generate_stats.Rmd
Expecting mean ~0, and upper/lower ~+/- 0.2
```{r permutation}
sd_threshold <- readRDS(file = here::here("data", "sd_threshold.rds"))
achilles_lower <- readRDS(file = here::here("data", "achilles_lower.rds"))
achilles_upper <- readRDS(file = here::here("data", "achilles_upper.rds"))
mean_virtual_achilles <- readRDS(file = here::here("data", "mean_virtual_achilles.rds"))
sd_virtual_achilles <- readRDS(file = here::here("data", "sd_virtual_achilles.rds"))
```

#how long to set-up this code for searches?
```{r}
end_time <- Sys.time()
time_taken <- round(as.duration(start_time %--% end_time)/dminutes(1), digits = 1)
print(time_taken)
```
Approximate time to run was `r time_taken` minutes.

#print Session information for provenance and reproducibility
```{r}
utils:::print.sessionInfo(sessionInfo()[-8]) 
#You can remove an item from sessionInfo(), which is a list with a class attribute, by printing the resulting object omitting one of the list items (omitted list of packages installed, but not loaded)

```

#beep
```{r}
beep(sound = 8) #because mario is awesome
```

#gene groups
```{r import_targets, eval=FALSE, include=FALSE}
nhr <- read_excel(here::here("data", "table1.xlsx"), col_names = FALSE) %>% 
  clean_names() %>% 
  pull() %>% 
  str_trim(side = c("both"))

#nhr_short <- nhr[!nhr %in% c("NR2E3", "AR", "NR0B1")] #chokes if I have columns that are not in dataset; need better way to skip over missing values #done; fixed with if/esle
#NHR_test <- NHR[NHR %in% c("THRA", "THRB")]
#nhr_trim <- nhr[48:48]
#str_which(nhr, "NR0B2")

slc <- id %>% 
  filter(str_detect(gene, "SLC")) %>% 
  pull(gene)

#slc <- str_subset(names(achilles_cor), "SLC")
#slc_test <- slc[1:10]
#str_which(slc, "SLC6A19")
#slc_trim <- slc[339:363]

SIRT <- c("SIRT1", "SIRT2", "SIRT3", "SIRT4", "SIRT5", "SIRT6", "SIRT7")
target_vec <- c("CPT1A", "CPT1B", "CPT1C", "CPT2", "SLC25A20")
hdac <- str_subset(names(achilles_cor), "HDAC")

mitocarta_full <- read_feather(here::here("data", "mitocarta.feather"))
mitocarta <- mitocarta_full %>% pull(Symbol)

transporter <- read_xlsx(here::here("data", "Transporter_genes_list.xlsx"), col_names = TRUE) %>% 
  clean_names() %>% 
  pull(gene)

#guest projects
trub <- c("FASN", "HMGCR", "HMGCS1")
transferase <- id %>% filter(str_detect(protein_name, "transferase")) %>% pull(gene)
acyltransferase <- id %>% filter(str_detect(protein_name, "acyltransferase")) %>% pull(gene)
deacylase <- id %>% filter(str_detect(protein_name, "deacetylase") | str_detect(protein_name, "deacylase")) %>% pull(gene)
abraham <- c("KCNK3", "KCNK9", "KCNK2")
angiotensin <- id %>% filter(str_detect(protein_name, "angiotensin")) %>% pull(gene)
myc <- id %>% filter(str_detect(gene, "MYC")) %>% pull(gene)
koves <- c("MLYCD")
klhl <- id %>% filter(str_detect(gene, "KLHL")) %>% pull(gene)
boyce <- c("GALE", "OGA", "OGT", "KEAP1", "GAN", "ENC1", klhl)
keap <- c("GAN", "ENC1")
bcaa <- c("ABAT", "ACAA1", "ACAA2", "ACAD8", "ACADM", "ACADS", "ACADSB", "ALDH1B1", "ALDH2", "ALDH3A2", "ALDH6A1", "ALDH7A1", "ALDH9A1", "AOX1", "AUH", "BCAT1", "BCAT2", "BCKDHA", "BCKDHB", "BCKDK", "DBT", "DLD", "ECHS1", "EHHADH", "HADH", "HADHA", "HADHB", "HIBADH", "HIBCH", "HMGCL", "HMGCL", "HMGCS2", "HSD17B10", "IVD", "MCCC1", "MCCC2", "MCEE", "MUT", "PCCA", "PCCB")
bcaa_trim <- c("MUT", "PCCA", "PCCB")
creatine <- c("GATM", "GAMT", "SLC6A8", "CKB", "CKM", "CKMT1B", "CKMT2")
glycolysis <- c("ALDOA", "BPGM", "ENO1", "ENO2", "GAPDH", "GPI", "HK1", "HK2", "HKDC1", "PFKL", "PFKM", "PGAM1", "PGAM2", "PGAM4", "PGK1", "PKLR", "PKM", "TPI1")
cardiolipin <- c("CKMT1A", "CKMT1B", "CYCS", "NME4", "TAZ", "TMEM256-PLSCR3", "CRLS1", "PGS1", "PTPMT1", "TAMM41")
cholesterol <- c("ACAT2", "CYP51A1", "DHCR24", "DHCR7", "EBP", "FDFT1", "FDPS", "GGPS1", "HMGCR", "HMGCS1", "HSD17B7", "IDI1", "LBR", "LIPA", "LSS", "MSMO1", "MVD", "MVK", "NSDHL", "PMVK", "SC5D", "SOAT1", "SQLE", "TM7SF2")
tca <- c("ACLY", "ACO1", "ACO2", "CS", "DLAT", "DLD", "DLST", "FH", "IDH1", "IDH2", "IDH3A", "IDH3B", "IDH3G", "MDH1", "MDH2", "MPC1", "OGDH", "OGDHL", "PC", "PCK1", "PCK2", "PDHA1", "PDHA2", "PDHB", "SDHA", "SDHB", "SDHC", "SDHD", "SUCLA2", "SUCLG1", "SUCLG2")
purine <- c("ADSL", "ADSS1", "ADSS2", "AK1", "AK2", "AK3", "AK4", "AK5", "ATIC", "GART", "GMPS", "IMPDH1", "IMPDH2", "PAICS", "PFAS", "PPAT", "PRPS1", "PRPS1L1", "PRPS2", "TAF9")
pyrimidine <- c("CAD", "CMPK1", "CMPK2", "CTPS1", "CTPS2", "DHODH", "NME1", "NME2", "NME3", "NME4", "NME6", "UMPS")
ampk <- c("CAB39", "CALM1", "CALM2", "CALM3", "CAMKK1", "CAMKK2", "PPM1E", "PPM1F", "PPP1CA", "PPP1CB", "PPP1R3D", "PPP2CA", "PRKAA1", "PRKAA2", "PRKAB1", "PRKAB2", "PRKAG1", "PRKAG2", "PRKAG3", "STK11", "STRADA")
ampk_trim <- c("PRKAA2", "PRKAB1", "PRKAB2", "PRKAG1", "PRKAG2", "PRKAG3", "STK11", "STRADA")
fao <- c("ABCD1", "ABCD2", "ACAA2", "ACAD9", "ACADL", "ACADM", "ACADS", "ACADVL", "ACAT1", "ACSL1", "ACSL3", "ACSL4", "ACSL5", "ACSL6", "ACSM1", "ACSM2A", "ACSM2B", "ACSM3", "ACSM4", "ACSS1", "ACSS2", "ACSS3", "CPT1A", "CPT1B", "CPT2", "CRAT", "CROT", "ECHS1", "ETFA", "ETFB", "ETFDH", "HADH", "HADHA", "HADHB", "HSD17B10", "PEX11G", "PEX13", "PEX14", "SLC25A20")
glycogen <- c("GBE1", "GCK", "GSK3A", "GSK3B", "GYG1", "GYS1", "GYS2", "PGM1", "PGM2", "PGM3", "PPP1CA", "PPP1CB", "PPP1CC", "PPP1R3A", "PPP1R3B", "PPP1R3C", "PPP1R3D", "PPP1R3E", "UGP2", "AGL", "CALM1", "CALM2", "CALM3", "G6PC", "PHKA1", "PHKA2", "PHKB", "PHKG1", "PHKG2", "PYGB", "PYGL")
fas <- c("ACACA", "ACACB", "ACLY", "FASN", "MCAT", "OLAH", "OXSM", "ACAA2", "ECHS1", "HADHA", "HADHB", "HSD17B10", "MECR", "PPT1")
urea <- c("ARG1", "ARG2", "ASL", "ASS1", "CPS1", "GLS2", "GLUD1", "GOT2", "GPT", "OTC", "SLC1A4", "SLC1A5", "SLC25A12", "SLC25A15")
ashley <- c("NMNAT1", "NMNAT2", "NMNAT3", "PNP")
tong <- c("MBOAT4")
herman <- c("MET", "HGF", "HGFAC", "REEP3", "PBX4", "KHK", "APOA5", "MLXIPL", "TM6SF2", "GALNT2", "MSL2", "TMEM184B")
hypoxia <- c("HIF1A", "ARNT", "EPAS1", "ARNT2", "HIF3A", "ARNT3", "EGLN1", "EGLN2", "EGLN3", "VHL", "HIF1AN")
counter <- c("RALA", "KRAS", "HRAS", "NRAS", "RAF1", "BRAF", "RRAS", "AURKA", "AURKB", "AURKC")
didier <- c("NPAS4", "ETV2", "EGFL7", "EMILIN3")
eoin <- c("EP300", "BRD4", "BRD2", "ACHY")
macalpine <- c("ORC1", "ORC2", "ORC3", "ORC4", "ORC5", "ORC6", "INO80", "SMARCE1", "SMARCD3", "SMARCA2", "SMARCA4", "SMARCB1", "SMARCA2", "SMARCD1", "SMARCC1", "SMARCC2" ,"CDC6")
alvarez <- c("TXN1", "TXN2", "PPAT", "PGD", "TXNRD1")
derek <- c("TAZ", "NOTCH1", "GATA1")
ferroptosis <- c("GPX4", "SLC7A11", "SLC3A2", "IREB2", "PHKG2", "HSPB1", "APP")
kimple <- c("GNAZ", "PTGER3")
scherer <- c("ADIPOQ", "ADIPOR1", "ADIPOR2", "LEP", "RETN", "RBP4", "CFD", "FGF21", "COL6A3")
silver <- c("MFSD8", "LAMP1", "MFSD1", "MFSD12", "FIT2")
#fav_gene <- c("LAMP1")
peter <- c("NDUFS4", "SURF1", "PMPCB", "COX10")
set <- c("SETD1A", "SETD2")
nih <- c("RFC4", "VPS18", "SMDT1")
acsf <-c("ACSF3")
gcn <- c("BLOC1S1")
hsp <- c("HSPD1")
pfk <- c("PFKFB1", "PFKFB2", "PFKFB3", "PFKFB4")
ben <- c("TMEM208", "CELSR1", "C10orf105")
jens <- c("GCGR", "GIPR")
mette <- c("ADGRA3", "GPR183") #GPR125
wood <- c("MTOR", "ETV4")
```

#loop to generate complete reports with enrichr_loop fun()
```{r reports_complete, eval=FALSE, warning=FALSE, include=FALSE}
gene_group <- wood

focused_lib <- c("Achilles_fitness_decrease", "Achilles_fitness_increase", "Aging_Perturbations_from_GEO_down", "Aging_Perturbations_from_GEO_up", "Allen_Brain_Atlas_down", "Allen_Brain_Atlas_up", "ARCHS4_Cell-lines", "ARCHS4_IDG_Coexp", "ARCHS4_Kinases_Coexp", "ARCHS4_TFs_Coexp", "ARCHS4_Tissues", "BioCarta_2016", "BioPlex_2017", "Cancer_Cell_Line_Encyclopedia", "ChEA_2016", "Chromosome_Location_hg19", "CORUM", "Data_Acquisition_Method_Most_Popular_Genes", "Disease_Perturbations_from_GEO_down", "Disease_Perturbations_from_GEO_up", "Disease_Signatures_from_GEO_up_2014", "Drug_Perturbations_from_GEO_down", "Drug_Perturbations_from_GEO_up", "DrugMatrix", "DSigDB", "ENCODE_and_ChEA_Consensus_TFs_from_ChIP-X", "ENCODE_Histone_Modifications_2015", "ENCODE_TF_ChIP-seq_2015", "Enrichr_Libraries_Most_Popular_Genes", "Enrichr_Submissions_TF-Gene_Coocurrence", "Epigenomics_Roadmap_HM_ChIP-seq", "ESCAPE", "GeneSigDB", "GO_Biological_Process_2018", "GO_Cellular_Component_2018", "GO_Molecular_Function_2018", "GTEx_Tissue_Sample_Gene_Expression_Profiles_down", "GTEx_Tissue_Sample_Gene_Expression_Profiles_up", "GWAS_Catalog_2019", "HMDB_Metabolites", "HomoloGene", "Human_Gene_Atlas", "Human_Phenotype_Ontology", "HumanCyc_2015", "HumanCyc_2016", "huMAP", "InterPro_Domains_2019", "Jensen_COMPARTMENTS", "Jensen_DISEASES", "Jensen_TISSUES", "KEA_2015", "KEGG_2019_Human", "KEGG_2019_Mouse", "Kinase_Perturbations_from_GEO_down", "Kinase_Perturbations_from_GEO_up", "Ligand_Perturbations_from_GEO_down", "Ligand_Perturbations_from_GEO_up", "LINCS_L1000_Chem_Pert_down", "LINCS_L1000_Chem_Pert_up", "LINCS_L1000_Kinase_Perturbations_down", "LINCS_L1000_Kinase_Perturbations_up", "LINCS_L1000_Ligand_Perturbations_down", "LINCS_L1000_Ligand_Perturbations_up", "MCF7_Perturbations_from_GEO_down", "MCF7_Perturbations_from_GEO_up", "MGI_Mammalian_Phenotype_Level_4_2019", "Microbe_Perturbations_from_GEO_down", "Microbe_Perturbations_from_GEO_up", "miRTarBase_2017", "Mouse_Gene_Atlas", "MSigDB_Computational", "MSigDB_Oncogenic_Signatures", "NCI-60_Cancer_Cell_Lines", "NURSA_Human_Endogenous_Complexome", "Old_CMAP_down", "Old_CMAP_up", "OMIM_Disease", "OMIM_Expanded", "Panther_2016", "Pfam_Domains_2019", "Pfam_InterPro_Domains", "Phosphatase_Substrates_from_DEPOD", "PPI_Hub_Proteins", "Rare_Diseases_AutoRIF_ARCHS4_Predictions", "Rare_Diseases_AutoRIF_Gene_Lists", "Rare_Diseases_GeneRIF_ARCHS4_Predictions", "Rare_Diseases_GeneRIF_Gene_Lists", "Reactome_2016", "RNA-Seq_Disease_Gene_and_Drug_Signatures_from_GEO", "SILAC_Phosphoproteomics", "Single_Gene_Perturbations_from_GEO_down", "Single_Gene_Perturbations_from_GEO_up", "SubCell_BarCode", "SysMyo_Muscle_Gene_Sets", "TargetScan_microRNA_2017", "TF_Perturbations_Followed_by_Expression", "TF-LOF_Expression_from_GEO", "Tissue_Protein_Expression_from_Human_Proteome_Map", "Tissue_Protein_Expression_from_ProteomicsDB", "Transcription_Factor_PPIs", "TRANSFAC_and_JASPAR_PWMs", "TRRUST_Transcription_Factors_2019", "UK_Biobank_GWAS", "Virus_Perturbations_from_GEO_down", "Virus_Perturbations_from_GEO_up", "VirusMINT", "WikiPathways_2019_Human", "WikiPathways_2019_Mouse")

for (fav_gene in gene_group) {
  if(fav_gene %in% names(achilles_cor) == 1){ #this code checks to see if the gene is in the analysis, and if not, skips
    #summary
    fav_gene_summary <- gene_summary %>% 
      filter(approved_symbol == fav_gene)
    
    #dep tables
    dep_top <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(desc(.[[2]])) %>% #use column index
      filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
      rename(approved_symbol = rowname) %>% 
      left_join(gene_summary, by = "approved_symbol") %>% 
      select(approved_symbol, approved_name, fav_gene) %>% 
      rename(gene = approved_symbol, name = approved_name, r2 = fav_gene)
    
    dep_bottom <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(.[[2]]) %>% #use column index
      filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
      rename(approved_symbol = rowname) %>% 
      left_join(gene_summary, by = "approved_symbol") %>% 
      select(approved_symbol, approved_name, fav_gene) %>% 
      rename(gene = approved_symbol, name = approved_name, r2 = fav_gene)
    #this is to get neg correlators
      
    #pathway enrichment analyses
    flat_top_complete <- dep_top %>%
       pull("gene") %>% 
       c(fav_gene, .) %>% 
       enrichr_loop(., focused_lib)
     
     #bottom 
     flat_bottom_complete <- dep_bottom %>%
       pull("gene") %>% 
       enrichr_loop(., focused_lib)

    #plot setup
    target_achilles <- achilles_long %>% 
      filter(gene == fav_gene) %>% 
      left_join(expression_join, by = "X1") %>% 
      select(cell_line, lineage, dep_score)
    
    target_achilles_top <- target_achilles %>% 
      top_frac(dep_score, n = 0.01)
    
    target_achilles_bottom <- target_achilles %>% 
      top_frac(dep_score, n = -0.01) %>% 
      arrange(dep_score)
    
    #plot1
    dep_plot1 <- ggplot(target_achilles) +
      geom_histogram(aes(x = dep_score), binwidth = 0.25, color = "lightgray") +
      labs(x = "Dependency Score (binned)") + 
      theme_light()
    
    #plot2
    dep_plot2 <- ggplot(target_achilles) +
      geom_point(aes(x = fct_rev(fct_reorder(target_achilles$cell_line, target_achilles$dep_score, .desc = TRUE)), y = dep_score)) +
      labs(x = "Cell Lines", y = "Dependency Score") +
      geom_hline(yintercept = mean_virtual_achilles) +
      geom_hline(yintercept = achilles_upper, linetype="dashed") +
      geom_hline(yintercept = achilles_lower, linetype="dashed") +
      geom_hline(yintercept = 0) +
      geom_point(data = target_achilles_top, aes(x = cell_line, y = dep_score), color = "red") +
      geom_point(data = target_achilles_bottom, aes(x = cell_line, y = dep_score), color = "red") +
      theme_light() +
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) + # axis.title.x=element_blank()
      NULL
    
    #render output
    render("report_depmap_complete.rmd", output_dir = here::here("results"), output_file = paste0(fav_gene, '_depmap.pdf'))
  } else {
    fav_gene_summary <- gene_summary %>% 
      filter(approved_symbol == fav_gene)
    
    #render output
    render("report_dummy_depmap.rmd", output_dir = here::here("results"), output_file = paste0(fav_gene, '_depmap.pdf'))
  }
}

#beep(sound = 8) 
#because mario is awesome
```

#cover page
```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
#render output
render("report_cover.rmd", output_dir = here::here("results"), output_file = "cover.pdf")
```


