---
title: "Dep Map correlation Network analysis"
output:
  pdf_document:
    toc: TRUE
params:
  release: "19Q3"
---
Overall goal of this project is to generate a network from the Dep Map data. Two goals:
1. Generate network of most connected/highest r2 values
2. Make networks of specific gene groups (this will then highlight the clusters within the gene groups, and then identify unexpected genes in known biological processes?)

##Load libraries
```{r load_block, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(feather)
library(corrr)
library(purrr)
library(tidygraph)
library(igraph)
library(ggraph)
library(ggforce)
library(networkD3)
library(viridis)
library(enrichR)
library(glue)
library(moderndive)
library(rmarkdown)
library(gridExtra)
library(knitr)
library(pander)
library(vroom)
library(beepr) #long analysis; get some coffee, and comeback when ready

#clear environment
#rm(list=ls()) 
```

##define functions
```{r}
make_graph <- function(gene_symbol, threshold = 10) {
  dep_network <- tibble()
  
 #find top and bottom correlations for fav_gene
  dep_top <- make_top_table(gene_symbol) %>% 
    slice(1:threshold)
  
  dep_bottom <- make_bottom_table(gene_symbol) %>% 
    slice(1:threshold) #limit for visualization?
  
  #this takes the genes from the top and bottom, and pulls them to feed them into a for loop
  related_genes <- dep_top %>% 
    bind_rows(dep_bottom) %>% 
    dplyr::pull(gene)
  
  #this loop will take each gene, and get their top and bottom correlations, and build a df containing the top n number of genes for each gene
  for (i in related_genes){
    message("Getting correlations related to ", gene_symbol, ", including ", i)
    dep_top_related <- achilles_cor %>% 
      focus(i) %>% 
      arrange(desc(.[[2]])) %>% #use column index
      filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
      mutate(x = i) %>% 
      rename(y = rowname, r2 = i) %>% 
      select(x, y, r2) %>% 
      slice(1:threshold) #limit for visualization?
    
    dep_bottom_related <- achilles_cor %>% 
      focus(i) %>% 
      arrange(.[[2]]) %>% #use column index
      filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
      mutate(x = i) %>% 
      rename(y = rowname, r2 = i) %>% 
      select(x, y, r2) %>% 
      slice(1:threshold) #limit for visualization?
    
    #each temp object is bound together, and then bound to the final df for graphing
    dep_related <- dep_top_related %>% 
      bind_rows(dep_bottom_related)
    
    dep_network <- dep_network %>% 
      bind_rows(dep_related)
  }
  
  simpleNetwork(dep_network, width="100%")
}
```

##import
```{r import}
##BROAD
load(file=here::here("data", paste0(params$release, "_achilles.RData")))
load(file=here::here("data", paste0(params$release, "_achilles_cor.RData")))
load(file=here::here("data", "expression_join.RData"))

class(achilles_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df

achilles_long <- achilles %>% gather("gene", "dep_score", -X1)
achilles_cor_long <- achilles_cor %>% 
  stretch() #observations across 3 variables (x, y, r); longest step, but cannot get cor_long.feather to load properly

load(here::here("data", "gene_summary.RData"))
```

#make some small dfs for testing
```{r}
achilles_cor_small <- achilles_cor %>% slice(1:1000) %>% select(1:1000) #to optimize code
achilles_long_small <- achilles_long %>% sample_n(1000)

```

#stats; import saved files generated in depmap_generate_stats.Rmd
Expecting mean ~0, and upper/lower ~+/- 0.2
```{r permutation}
sd_threshold <- readRDS(file = here::here("data", "sd_threshold.rds"))
achilles_lower <- readRDS(file = here::here("data", "achilles_lower.rds"))
achilles_upper <- readRDS(file = here::here("data", "achilles_upper.rds"))
mean_virtual_achilles <- readRDS(file = here::here("data", "mean_virtual_achilles.rds"))
sd_virtual_achilles <- readRDS(file = here::here("data", "sd_virtual_achilles.rds"))
```

#single gene
This is a test case for a single gene. Takes a single input, and then generates top and bottom correlated genes. Finally, binds them together to make a graph structure.
```{r}
fav_gene <- c("TP53")
n <- 10
#find top and bottom correlations for fav_gene
dep_top <- achilles_cor %>% 
  focus(fav_gene) %>% 
  arrange(desc(.[[2]])) %>% #use column index
  filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
  mutate(x = fav_gene) %>% 
  rename(y = rowname, r2 = fav_gene) %>% 
  select(x, y, r2) %>% 
  slice(1:n) #limit for visualization?

dep_bottom <- achilles_cor %>% 
  focus(fav_gene) %>% 
  arrange(.[[2]]) %>% #use column index
  filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
  mutate(x = fav_gene) %>% 
  rename(y = rowname, r2 = fav_gene) %>% 
  select(x, y, r2) %>% 
  slice(1:n) #limit for visualization?

#last steps of above pipes are providing x, y, and r2 for graphing; this binds them together
dep_network <- dep_top %>% 
  bind_rows(dep_bottom)
```

#single gene and related genes
This builds on the code above, and creates a for-loop to generate a network graph of a single gene. It takes a gene, finds the n top and bottom correlated genes (called related genes), and then pulls the correlations of the related genes to builds up a df. 
```{r}
fav_gene <- c("TP53")
n <- 10
dep_network <- tibble(
  x = character(), 
  y = character(),
  r2 = numeric()
)

#find top and bottom correlations for fav_gene
dep_top <- achilles_cor %>% 
  focus(fav_gene) %>% 
  arrange(desc(.[[2]])) %>% #use column index
  filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
  mutate(x = fav_gene) %>% 
  rename(y = rowname, r2 = fav_gene) %>% 
  select(x, y, r2) %>% 
  slice(1:n) #limit for visualization?

dep_bottom <- achilles_cor %>% 
  focus(fav_gene) %>% 
  arrange(.[[2]]) %>% #use column index
  filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
  mutate(x = fav_gene) %>% 
  rename(y = rowname, r2 = fav_gene) %>% 
  select(x, y, r2) %>% 
  slice(1:n) #limit for visualization?

#this takes the genes from the top and bottom, and pulls them to feed them into a for loop
related_genes <- dep_top %>% 
  bind_rows(dep_bottom) %>% 
  dplyr::pull(y) #uncomment this

#this loop will take each gene, and get their top and bottom correlations, and build a df containing the top n number of genes for each gene
for (i in related_genes){
  message("Getting correlations related to ", fav_gene, ", including ", i)
  dep_top_related <- achilles_cor %>% 
    focus(i) %>% 
    arrange(desc(.[[2]])) %>% #use column index
    filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
    mutate(x = i) %>% 
    rename(y = rowname, r2 = i) %>% 
    select(x, y, r2) %>% 
    slice(1:n) #limit for visualization?
  
  dep_bottom_related <- achilles_cor %>% 
    focus(i) %>% 
    arrange(.[[2]]) %>% #use column index
    filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
    mutate(x = i) %>% 
    rename(y = rowname, r2 = i) %>% 
    select(x, y, r2) %>% 
    slice(1:n) #limit for visualization?
  
  #each temp object is bound together, and then bound to the final df for graphing
  dep_related <- dep_top_related %>% 
    bind_rows(dep_bottom_related)
  
  dep_network <- dep_network %>% 
    bind_rows(dep_related)
}
```

#build graph from gene + related genes
```{r}
#minimum degree
m <- 3

#https://www.r-bloggers.com/another-game-of-thrones-network-analysis-this-time-with-tidygraph-and-ggraph/
graph_network <- as_tbl_graph(dep_network)
graph_network <- graph_network %>% 
  activate(nodes) %>% #active by default, add centrality measures to nodes
  mutate(degree = degree(graph_network), #The degree of a node is the number of its adjacent edges.
         strength = strength(graph_network), #Summing up the edge weights of the adjacent edges for each vertex (same as degree, if not weighted)
         betweenness = betweenness(graph_network), #quantifies how often a node lies on the shortest path between other nodes (measure of ties)
         closeness = closeness(graph_network),
         neighbors = centrality_degree(), 
         center = node_is_center(),
         keyplayer = node_is_keyplayer(k = 10),
         query = (name %in% related_genes), 
         n_rank_trv = node_rank_traveller()) %>%  #how close a node is to all other nodes in the network in terms of shortest path distance
  arrange(desc(degree)) %>% 
  #arrange(n_rank_trv) %>% 
  activate(edges) %>% 
  filter(!edge_is_multiple()) %>%  # remove multiple edges
  mutate(r2_bin = if_else(r2 > 0, "positive", "negative"), 
         centrality_e = centrality_edge_betweenness(directed = FALSE))

nodes <- graph_network %>% 
  activate(nodes) %>% # %N>%
  as_tibble()

edges <- graph_network %>% 
  activate(edges) %>% # %E>%
  as_tibble()
 #graph_network
  
graph_network %>% 
  ggraph(layout = 'kk', weights = abs(r2)) +
  geom_edge_link(aes(filter = centrality_e > 1000, color = r2_bin), alpha = 0.5) +
  #scale_edge_color_viridis(discrete = TRUE, option = "magma") +
  scale_edge_colour_brewer(palette = "Set1", direction = -1) +
  geom_node_point(aes(color = query), size = 6) +
  scale_color_manual(values = c("yellow", "gray")) +
  geom_node_text (aes(label = name), size = 3, repel = FALSE) +
  theme_graph()

graph_network %>% 
  ggraph(layout = 'kk', weights = abs(r2)) +
  geom_edge_link(aes(filter = centrality_e > 1000, color = r2_bin), alpha = 0.5) +
  #scale_edge_color_viridis(discrete = TRUE, option = "magma") +
  scale_edge_colour_brewer(palette = "Set1", direction = -1) +
  geom_node_point(aes(filter = betweenness > 1, color = query), size = 6) +
  scale_color_manual(values = c("yellow", "gray")) +
  geom_node_text (aes(filter = betweenness > 1, label = name), size = 3, repel = FALSE) +
  theme_graph()

#https://www.r-graph-gallery.com/network-interactive.html
simpleNetwork(dep_network, height="500px", width="500px")
```

#build forceNetwork
kaggle; https://www.kaggle.com/paulc91/visualise-flows-with-r-d3
```{r}
build_graph <- function(dep_network = dep_network, deg = 2) {
  #setup graph 
  graph_network <- tidygraph::as_tbl_graph(dep_network)
  nodes <-  as_tibble(graph_network) %>% 
    rowid_to_column("id") %>% 
    mutate(degree = degree(graph_network), 
           group = 1) %>% 
    arrange(desc(degree))
  
  links <- graph_network %>% 
    activate(edges) %>% # %E>%
    as_tibble()
  
  # determine the nodes that have at least the minimum degree
  nodes_filtered <- nodes %>%
    filter(degree >= deg) %>%  #input$degree
    as.data.frame
  
  # filter the edge list to contain only links to or from the nodes that have the minimum or more degree
  links_filtered <- links %>% 
    filter(to %in% nodes_filtered$id & from %in% nodes_filtered$id) %>% 
    as.data.frame
  
  # re-adjust the from and to values to reflect the new positions of nodes in the filtered nodes list
  links_filtered$from <- match(links_filtered$from, nodes_filtered$id) - 1
  links_filtered$to <- match(links_filtered$to, nodes_filtered$id) - 1
  
  forceNetwork(Links = links_filtered, Nodes = nodes_filtered, Source = "from", Target ="to", NodeID = "name", Group = "group", zoom = TRUE, bounded = TRUE, opacityNoHover = 100)
}
build_graph(dep_network, 2)
build_graph(dep_network, 10)
```


#gene lists
```{r}
single <- c("TP53")
triple <- c("TP53", "MDM2", "MDM4")
glycolysis <- c("ALDOA", "BPGM", "ENO1", "ENO2", "GAPDH", "GPI", "HK1", "HK2", "HKDC1", "PFKL", "PFKM", "PGAM1", "PGAM2", "PGAM4", "PGK1", "PKLR", "PKM", "TPI1")
tca <- c("ACLY", "ACO1", "ACO2", "CS", "DLAT", "DLD", "DLST", "FH", "IDH1", "IDH2", "IDH3A", "IDH3B", "IDH3G", "MDH1", "MDH2", "MPC1", "OGDH", "OGDHL", "PC", "PCK1", "PCK2", "PDHA1", "PDHA2", "PDHB", "SDHA", "SDHB", "SDHC", "SDHD", "SUCLA2", "SUCLG1", "SUCLG2")
purine <- c("ADSL", "ADSS1", "ADSS2", "AK1", "AK2", "AK3", "AK4", "AK5", "ATIC", "GART", "GMPS", "IMPDH1", "IMPDH2", "PAICS", "PFAS", "PPAT", "PRPS1", "PRPS1L1", "PRPS2", "TAF9")
pyrimidine <- c("CAD", "CMPK1", "CMPK2", "CTPS1", "CTPS2", "DHODH", "NME1", "NME2", "NME3", "NME4", "NME6", "UMPS")
```

#gene list
This builds further, and creates a df from a list of genes. Importantly, it omits the step above to find related genes, because presumably the gene list is related (also, it'd generate a BIG recursive loop)
```{r}
n <- 25
gene_list <- pyrimidine
dep_network <- tibble()
for (fav_gene in gene_list) {
  message("Getting correlations for ", fav_gene)
  if(fav_gene %in% names(achilles_cor) == 1){ #this code checks to see if the gene is in the analysis, and if not, skips
    #find top and bottom correlations for fav_gene
    dep_top <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(desc(.[[2]])) %>% #use column index
      filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
      mutate(x = fav_gene) %>% 
      rename(y = rowname, r2 = fav_gene) %>% 
      select(x, y, r2) %>% 
      slice(1:n) #limit for visualization?
    
    #dep_bottom <- achilles_cor %>% 
    #  focus(fav_gene) %>% 
    #  arrange(.[[2]]) %>% #use column index
    #  filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
    #  mutate(x = fav_gene) %>% 
    #  rename(y = rowname, r2 = fav_gene) %>% 
    #  select(x, y, r2) %>% 
    #  slice(1:n) #limit for visualization?
  } else {
    dep_top <- tibble(
      x = character(), 
      y= character(),
      r2 = numeric())
  }
  #last steps of above pipes are providing x, y, and r2 for graphing; this binds them together
  dep_network <- dep_network %>% 
    bind_rows(dep_top)
}
```

#graph
Thinking about graph_netwrok as two tibbles inside a larger table graph, it became evident that dplyr needs a way to know which of the two tables (nodes or edges) to perform the transformations on. In tidygraph, this is done using the activate() function. To showcase this, the nodes table will be “activated” in order to add two new string variables derived from name.
```{r}
#minimum degree
m <- 3

graph_network <- as_tbl_graph(dep_network) %>% 
  activate(nodes) %>% #active by default, add centrality measures to nodes
  mutate(degree = degree(graph_network), #The degree of a node is the number of its adjacent edges.
         strength = strength(graph_network), #Summing up the edge weights of the adjacent edges for each vertex (same as degree, if not weighted)
         betweenness = betweenness(graph_network), #quantifies how often a node lies on the shortest path between other nodes (measure of ties)
         closeness = closeness(graph_network),
         neighbors = centrality_degree(), 
         center = node_is_center(),
         keyplayer = node_is_keyplayer(k = 10),
         query = (name %in% related_genes), 
         n_rank_trv = node_rank_traveller()) %>%  #how close a node is to all other nodes in the network in terms of shortest path distance
  arrange(desc(degree)) %>% 
  #arrange(n_rank_trv) %>% 
  activate(edges) %>% 
  filter(!edge_is_multiple()) %>%  # remove multiple edges
  mutate(r2_bin = if_else(r2 > 0, "positive", "negative"), 
         centrality_e = centrality_edge_betweenness(directed = FALSE))

nodes <- graph_network %>% 
  activate(nodes) %>% # %N>%
  as_tibble()

edges <- graph_network %>% 
  activate(edges) %>% # %E>%
  as_tibble()
 #graph_network
  
graph_network %>% 
  ggraph(layout = 'kk', weights = abs(r2)) +
  geom_edge_link(aes(filter = centrality_e > 100, color = r2_bin), alpha = 0.5) +
  #scale_edge_color_viridis(discrete = TRUE, option = "magma") +
  scale_edge_colour_brewer(palette = "Set1", direction = -1) +
  geom_node_point(aes(filter = betweenness > 1, color = query), size = 6) +
  scale_color_manual(values = c("yellow", "gray")) +
  geom_node_text (aes(filter = betweenness > 1, label = name), size = 3, repel = FALSE) +
  theme_graph()

```

#alt graphs
```{r eval = FALSE}
graph_network %>% 
  filter(degree > 1) %>% 
  ggraph('focus', focus = node_is_center()) + 
  ggforce::geom_circle(aes(x0 = 0, y0 = 0, r = r), data.frame(r = 1:3), colour = 'grey') + 
  geom_edge_link() + 
  geom_node_point(aes(color = query)) + 
  coord_fixed()


graph_network %>% 
  filter(degree > 1) %>% 
  ggraph('backbone') + 
  ggforce::geom_circle(aes(x0 = 0, y0 = 0, r = r), data.frame(r = 1:3), colour = 'grey') + 
  geom_edge_link() + 
  geom_node_point(aes(color = query)) + 
  coord_fixed()


#graph_network %>%
# ggraph() +
#  geom_edge_link(aes(start_cap = label_rect(node1.name),
#                     end_cap = label_rect(node2.name))) + 
#  geom_node_text(aes(label = name))


```


#MAKE GRAPHS PRETTY
https://drsimonj.svbtle.com/how-to-create-correlation-network-plots-with-corrr-and-ggraph
Basic approach #

Given a data frame d of numeric variables for which we want to plot the correlations in a network, here’s a basic approach:

*Create a tidy data frame of correlations*
tidy_cors <- d %>% 
  correlate() %>% 
  stretch()

*Convert correlations stronger than some value to an undirected graph object*
graph_cors <- tidy_cors %>% 
  filter(abs(r) > `VALUE_BETWEEN_0_AND_1`) %>% 
  graph_from_data_frame(directed = FALSE)

*Plot*
ggraph(graph_cors) +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph()
  
Check out: https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/
WGCNA: an R package for weighted correlation network analysis

https://www.data-imaginist.com/2017/announcing-ggraph/

```{r graphs, eval=FALSE, warning=FALSE, include=FALSE}
graph_cors <- achilles_cor_long %>% 
  filter(abs(r) > .7) %>%
  #filter(r  > achilles_upper | r < achilles_lower) %>% #+/- 3SD of mean distribution
  graph_from_data_frame(directed = FALSE)

ggraph(graph_cors, layout = "auto") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name)) +
  theme_graph()

ggraph(graph_cors, layout = "auto") +
  geom_edge_link(aes(edge_alpha = abs(r), edge_width = abs(r), color = r)) +
  guides(edge_alpha = "none", edge_width = "none") +
  scale_edge_colour_gradientn(limits = c(-1, 1), colors = c("blue", "red")) +
  geom_node_point(color = "black", size = 2, alpha = 0.5) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph() +
  labs(title = "Correlations between genetic dependencies")

#need to make adj network? 
#https://campus.datacamp.com/courses/network-science-in-r-a-tidy-approach/connection-patterns?ex=1

#need to reduce number of networks, and/or highlight networks. how can you facet based on a sub-network? visualize with facet_node(subnetwork_id)
#https://en.wikipedia.org/wiki/Adjacency_matrix
#https://www.data-imaginist.com/2017/announcing-ggraph/

```

