---
title: "Dep Map correlation Network analysis"
output:
  pdf_document:
    toc: TRUE
params:
  release: "19Q3"
---
Overall goal of this project is to generate a network from the Dep Map data. Two goals:
1. Generate network of most connected/highest r2 values
2. Make networks of specific gene groups (this will then highlight the clusters within the gene groups, and then identify unexpected genes in known biological processes?)

##Load libraries
```{r load_block, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(here)
library(janitor)
library(feather)
library(corrr)
library(purrr)
library(igraph)
library(ggraph)
library(viridis)
library(enrichR)
library(glue)
library(moderndive)
library(rmarkdown)
library(gridExtra)
library(knitr)
library(pander)
library(vroom)
library(beepr) #long analysis; get some coffee, and comeback when ready

#clear environment
#rm(list=ls()) 
```

##import
Import Broad Insititute and Sanger Depedencay data; added Broad cell line expression data (Aug 21, 2019).
```{r import}
##BROAD
achilles <- read_feather(here::here("data", paste0(params$release, "_achilles.feather")))
achilles_cor <- read_feather(here("data", paste0(params$release, "_achilles_cor.feather"))) 
class(achilles_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df
achilles_cor_long <- read_feather(here::here("data", paste0(params$release, "_achilles_cor_long.feather")))

#EXPRESSION(BROAD)
expression_id <- read_feather(here("data", paste0(params$release, "_expression_id.feather"))) 
expression_join <- expression_id %>% 
  rename(X1 = dep_map_id) %>% 
  select(X1, stripped_cell_line_name, lineage)

```

#make some small dfs for testing
```{r}
achilles_cor_small <- achilles_cor %>% slice(1:1000) %>% select(1:1000) #to optimize code
achilles_long_small <- achilles_long %>% sample_n(1000)

```

#Correlation network for top and bottom genes
```{r network}
#top
achilles_cor_top50 <- achilles_cor %>% 
  select(-rowname) %>% #select(one_of(target_vec)) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) %>% 
  top_n(50) 

pander(achilles_cor_top50)
  
achilles_cor_bottom50 <- achilles_cor %>% 
  select(-rowname) %>% #select(one_of(target_vec)) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>% 
  gather("gene", "min") %>% 
  arrange(min)  %>% 
  top_n(-50)

pander(achilles_cor_bottom50)

focus_50 <- c(achilles_cor_top50, achilles_cor_bottom50)

expression_cor_max %>% 
  top_n(50)

#several r2 = 1.0
```

#MAKE GRAPHS PRETTY
https://drsimonj.svbtle.com/how-to-create-correlation-network-plots-with-corrr-and-ggraph
Basic approach #

Given a data frame d of numeric variables for which we want to plot the correlations in a network, here’s a basic approach:

*Create a tidy data frame of correlations*
tidy_cors <- d %>% 
  correlate() %>% 
  stretch()

*Convert correlations stronger than some value to an undirected graph object*
graph_cors <- tidy_cors %>% 
  filter(abs(r) > `VALUE_BETWEEN_0_AND_1`) %>% 
  graph_from_data_frame(directed = FALSE)

*Plot*
ggraph(graph_cors) +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph()
  
Check out: https://horvath.genetics.ucla.edu/html/CoexpressionNetwork/Rpackages/WGCNA/
WGCNA: an R package for weighted correlation network analysis

https://www.data-imaginist.com/2017/announcing-ggraph/

```{r graphs, eval=FALSE, warning=FALSE, include=FALSE}
graph_cors <- achilles_cor_long %>% 
  filter(abs(r) > .7) %>%
  #filter(r  > achilles_upper | r < achilles_lower) %>% #+/- 3SD of mean distribution
  graph_from_data_frame(directed = FALSE)

ggraph(graph_cors, layout = "auto") +
  geom_edge_link() +
  geom_node_point() +
  geom_node_text(aes(label = name)) +
  theme_graph()

ggraph(graph_cors, layout = "auto") +
  geom_edge_link(aes(edge_alpha = abs(r), edge_width = abs(r), color = r)) +
  guides(edge_alpha = "none", edge_width = "none") +
  scale_edge_colour_gradientn(limits = c(-1, 1), colors = c("blue", "red")) +
  geom_node_point(color = "black", size = 2, alpha = 0.5) +
  geom_node_text(aes(label = name), repel = TRUE) +
  theme_graph() +
  labs(title = "Correlations between genetic dependencies")

#need to make adj network? 
#https://campus.datacamp.com/courses/network-science-in-r-a-tidy-approach/connection-patterns?ex=1

#need to reduce number of networks, and/or highlight networks. how can you facet based on a sub-network? visualize with facet_node(subnetwork_id)
#https://en.wikipedia.org/wiki/Adjacency_matrix
#https://www.data-imaginist.com/2017/announcing-ggraph/

```

