---
title: "Dep Map analysis pathway generator"
output:
  pdf_document:
    toc: TRUE
params:
  release: "19Q3"
---
This document loads the data generated in depmap_generate, and then attemps to generate pathway data for top and bottom correlated genes, generate the plots, and store them in three distince list-column dataframes:  

master_postive  
master_negative  
master_plots  

##Load libraries
```{r load_block, echo=TRUE, message=FALSE, warning=FALSE}
#not sure which exactly of these I need
library(tidyverse)
library(lubridate)
library(here)
library(readxl)
library(janitor)
library(feather)
library(corrr)
library(purrr)
library(viridis)
library(enrichR)
library(glue)
library(moderndive)
library(rmarkdown)
library(gridExtra)
library(knitr)
library(pander)
library(vroom)
library(beepr) #long analysis; get some coffee, and comeback when ready

#clear environment
#rm(list=ls()) 

#how long?
start_time <- Sys.time()
```

#define functions
```{r}
#pathway enrichment analysis loop function
enrichr_loop <- function(gene_list, databases){
  if(is_empty(gene_list)){
    flat_complete <- NULL
    return(flat_complete)
  } else {
    flat_complete <- as_tibble()
    for(lib in databases){
      enriched <- enrichr(gene_list, lib)
      
      flat <- flatten_dfc(enriched) %>% 
        mutate(enrichr = lib)
      
      flat_complete <- flat_complete %>% 
        bind_rows(flat)
    }
    flat_complete <- flat_complete %>% 
      arrange(Adjusted.P.value) 
      #select(enrichr, Term, Overlap)
    
    flat_complete$enrichr <- str_replace_all(flat_complete$enrichr, "\\_", " ")
    flat_complete$Term <- str_replace_all(flat_complete$Term, "\\_", " ")
    return(flat_complete)
  }
}
```

##import
```{r import}
##BROAD
achilles <- read_feather(here::here("data", paste0(params$release, "_achilles.feather")))
achilles_long <- achilles %>% gather("gene", "dep_score", -X1)
achilles_cor <- read_feather(here::here("data", paste0(params$release, "_achilles_cor.feather"))) 
class(achilles_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df

achilles_cor_long <- achilles_cor %>% 
  stretch() #observations across 3 variables (x, y, r); longest step, but cannot get cor_long.feather to load properly

#EXPRESSION(BROAD), need this to get cell line names
expression_id <- read_feather(here::here("data", paste0(params$release, "_expression_id.feather"))) 

expression_join <- expression_id %>% 
  rename(X1 = dep_map_id) %>% 
  select(X1, stripped_cell_line_name, lineage)

gene_summary <- read_feather(path = here::here("data", "gene_summary.feather"))
```

#Permutation tests
A permutation test involves permuting one or more variables in a data set before performing the test, in order to break any existing relationships and simulate the null hypothesis. In this case, we will break the relationshipe between gene-gene pairs and the correlation values. We can then compare the true statistic (mean correlation) to the generated distribution of null statistics (fake means). This will give a better idea of where to draw a threshold of a "signficiant correlation" for these analyses. Recalculate here for statistical threshold cutoffs.

```{r permutation, eval=FALSE}
  virtual_achilles <- achilles_cor_long %>% #achilles_cor_long already has all of the variables in a long format
  filter(!is.na(r)) %>%   
  rep_sample_n(size = 20000, reps = 1000) %>% #larger sample size, less error (but only 625 sets, and we're mimicking 1000?, but 310M combinations, so probably OK)
  group_by(replicate) %>% 
  summarize(mean = mean(r), max = max(r), min = min(r), sd = sd(r)) #how to handle + vs. - correlation?

#instead of filtering, could:
#summarize_all(list(max = max(r, na.rm = TRUE), 
#                     min = min(r, na.rm = TRUE),
#                     mean = mean(r, na.rm = TRUE),
#                     sd = sd(r, na.rm = TRUE))) 
#but threw and error

#The mean may not be the best measure of centrality due to its sensitivity to outliers. The example data used here seems asymmetric and it that case the median is a better measure of centrality.
#https://github.com/JoachimGoedhart/PlotsOfDifferences

ggplot(virtual_achilles, aes(x = sd)) +
  geom_histogram(color = "white") +
  labs(x = "SD r^2 value from virtual achilles", 
       title = "Distribution of mean r^2 values") 

mean_virtual_achilles <- mean(virtual_achilles$mean)
sd_virtual_achilles <- mean(virtual_achilles$sd)

sd_threshold <- 3

achilles_upper <- mean_virtual_achilles + sd_threshold*sd_virtual_achilles
achilles_lower <- mean_virtual_achilles - sd_threshold*sd_virtual_achilles


#ADD GENERATE AND RELOAD PERMUTATION DATA HERE.

#permutation_tests <- (list(mean_virtual_achilles, 
#                           sd_virtual_achilles, 
#                           sd_threshold,
#                           achilles_upper, 
#                           achilles_lower))
#save(permutation_tests, file = here::here("data", "permutation_tests.Rdata"))
#load(here::here("data", "permutation_tests.Rdata"))

saveRDS(sd_threshold, file = here::here("data", "sd_threshold.rds"))
saveRDS(achilles_lower, file = here::here("data", "achilles_lower.rds"))
saveRDS(achilles_upper, file = here::here("data", "achilles_upper.rds"))
saveRDS(mean_virtual_achilles, file = here::here("data", "mean_virtual_achilles.rds"))
saveRDS(sd_virtual_achilles, file = here::here("data", "sd_virtual_achilles.rds"))

#sd_threshold <- readRDS(file = here::here("data", "sd_threshold.rds"))
#rm(sd_threshold) test...it works
#achilles_lower <- readRDS(file = here::here("data", "achilles_lower.rds"))
#achilles_upper <- readRDS(file = here::here("data", "achilles_upper.rds"))
#mean_virtual_achilles <- readRDS(file = here::here("data", "mean_virtual_achilles.rds"))
#sd_virtual_achilles <- readRDS(file = here::here("data", "sd_virtual_achilles.rds"))
```

#setup containers
```{r eval=FALSE}
#running this overwrites any master tables in the environment
master_positive <- tibble(
  fav_gene = character(), 
  data = list()
)
master_negative <- tibble(
  fav_gene = character(), 
  data = list()
)
master_plots <- tibble(
  fav_gene = character(), 
  plot1 = list(), 
  plot2 = list()
)
master_top_table <- tibble(
  fav_gene = character(), 
  data = list()
)
master_bottom_table <- tibble(
  fav_gene = character(), 
  data = list()
)
```

#temp data groups 
```{r}
purine <- c("ADSL", "ADSS1", "ADSS2", "AK1", "AK2", "AK3", "AK4", "AK5", "ATIC", "GART", "GMPS", "IMPDH1", "IMPDH2", "PAICS", "PFAS", "PPAT", "PRPS1", "PRPS1L1", "PRPS2", "TAF9")
pyrimidine <- c("CAD", "CMPK1", "CMPK2", "CTPS1", "CTPS2", "DHODH", "NME1", "NME2", "NME3", "NME4", "NME6", "UMPS")

set.seed(2019)
sample <- sample(names(achilles), size = 10)
```

#put data into lists
```{r reports_complete, eval=FALSE, warning=FALSE, include=FALSE}
gene_group <- sample

focused_lib <- c("Achilles_fitness_decrease", "Achilles_fitness_increase", "Aging_Perturbations_from_GEO_down", "Aging_Perturbations_from_GEO_up", "Allen_Brain_Atlas_down", "Allen_Brain_Atlas_up", "ARCHS4_Cell-lines", "ARCHS4_IDG_Coexp", "ARCHS4_Kinases_Coexp", "ARCHS4_TFs_Coexp", "ARCHS4_Tissues", "BioCarta_2016", "BioPlex_2017", "Cancer_Cell_Line_Encyclopedia", "ChEA_2016", "Chromosome_Location_hg19", "CORUM", "Data_Acquisition_Method_Most_Popular_Genes", "Disease_Perturbations_from_GEO_down", "Disease_Perturbations_from_GEO_up", "Disease_Signatures_from_GEO_up_2014", "Drug_Perturbations_from_GEO_down", "Drug_Perturbations_from_GEO_up", "DrugMatrix", "DSigDB", "ENCODE_and_ChEA_Consensus_TFs_from_ChIP-X", "ENCODE_Histone_Modifications_2015", "ENCODE_TF_ChIP-seq_2015", "Enrichr_Libraries_Most_Popular_Genes", "Enrichr_Submissions_TF-Gene_Coocurrence", "Epigenomics_Roadmap_HM_ChIP-seq", "ESCAPE", "GeneSigDB", "GO_Biological_Process_2018", "GO_Cellular_Component_2018", "GO_Molecular_Function_2018", "GTEx_Tissue_Sample_Gene_Expression_Profiles_down", "GTEx_Tissue_Sample_Gene_Expression_Profiles_up", "GWAS_Catalog_2019", "HMDB_Metabolites", "HomoloGene", "Human_Gene_Atlas", "Human_Phenotype_Ontology", "HumanCyc_2015", "HumanCyc_2016", "huMAP", "InterPro_Domains_2019", "Jensen_COMPARTMENTS", "Jensen_DISEASES", "Jensen_TISSUES", "KEA_2015", "KEGG_2019_Human", "KEGG_2019_Mouse", "Kinase_Perturbations_from_GEO_down", "Kinase_Perturbations_from_GEO_up", "Ligand_Perturbations_from_GEO_down", "Ligand_Perturbations_from_GEO_up", "LINCS_L1000_Chem_Pert_down", "LINCS_L1000_Chem_Pert_up", "LINCS_L1000_Kinase_Perturbations_down", "LINCS_L1000_Kinase_Perturbations_up", "LINCS_L1000_Ligand_Perturbations_down", "LINCS_L1000_Ligand_Perturbations_up", "MCF7_Perturbations_from_GEO_down", "MCF7_Perturbations_from_GEO_up", "MGI_Mammalian_Phenotype_Level_4_2019", "Microbe_Perturbations_from_GEO_down", "Microbe_Perturbations_from_GEO_up", "miRTarBase_2017", "Mouse_Gene_Atlas", "MSigDB_Computational", "MSigDB_Oncogenic_Signatures", "NCI-60_Cancer_Cell_Lines", "NURSA_Human_Endogenous_Complexome", "Old_CMAP_down", "Old_CMAP_up", "OMIM_Disease", "OMIM_Expanded", "Panther_2016", "Pfam_Domains_2019", "Pfam_InterPro_Domains", "Phosphatase_Substrates_from_DEPOD", "PPI_Hub_Proteins", "Rare_Diseases_AutoRIF_ARCHS4_Predictions", "Rare_Diseases_AutoRIF_Gene_Lists", "Rare_Diseases_GeneRIF_ARCHS4_Predictions", "Rare_Diseases_GeneRIF_Gene_Lists", "Reactome_2016", "RNA-Seq_Disease_Gene_and_Drug_Signatures_from_GEO", "SILAC_Phosphoproteomics", "Single_Gene_Perturbations_from_GEO_down", "Single_Gene_Perturbations_from_GEO_up", "SubCell_BarCode", "SysMyo_Muscle_Gene_Sets", "TargetScan_microRNA_2017", "TF_Perturbations_Followed_by_Expression", "TF-LOF_Expression_from_GEO", "Tissue_Protein_Expression_from_Human_Proteome_Map", "Tissue_Protein_Expression_from_ProteomicsDB", "Transcription_Factor_PPIs", "TRANSFAC_and_JASPAR_PWMs", "TRRUST_Transcription_Factors_2019", "UK_Biobank_GWAS", "Virus_Perturbations_from_GEO_down", "Virus_Perturbations_from_GEO_up", "VirusMINT", "WikiPathways_2019_Human", "WikiPathways_2019_Mouse")

focused_lib_small <- c("Achilles_fitness_decrease", "Achilles_fitness_increase")

for (fav_gene in gene_group) {
  if(fav_gene %in% names(achilles_cor) == 1){ #this code checks to see if the gene is in the analysis, and if not, skips
    
    message("Top dep tables for ", fav_gene)
    dep_top <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(desc(.[[2]])) %>% #use column index
      filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
      rename(approved_symbol = rowname) %>% 
      left_join(gene_summary, by = "approved_symbol") %>% 
      select(approved_symbol, approved_name, fav_gene) %>% 
      rename(gene = approved_symbol, name = approved_name, r2 = fav_gene)
    
    top_table <- dep_top %>% 
      mutate(fav_gene = fav_gene) %>% 
      group_by(fav_gene) %>% 
      nest()
    
    master_top_table <- master_top_table %>% 
      bind_rows(top_table)
    
    message("Bottom dep tables for ", fav_gene)
    dep_bottom <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(.[[2]]) %>% #use column index
      filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
      rename(approved_symbol = rowname) %>% 
      left_join(gene_summary, by = "approved_symbol") %>% 
      select(approved_symbol, approved_name, fav_gene) %>% 
      rename(gene = approved_symbol, name = approved_name, r2 = fav_gene)
    
    bottom_table <- dep_bottom %>% 
      mutate(fav_gene = fav_gene) %>% 
      group_by(fav_gene) %>% 
      nest()
    
    master_bottom_table <- master_bottom_table %>% 
      bind_rows(bottom_table)
    
    message("Top pathway enrichment analyses for ", fav_gene)
    flat_top_complete <- dep_top %>%
      pull("gene") %>% 
      c(fav_gene, .) %>% 
      enrichr_loop(., focused_lib) %>%  #added small here
      top_n(100)
    
    positive <- flat_top_complete  %>% 
      mutate(fav_gene = fav_gene) %>% 
      group_by(fav_gene) %>% 
      nest()
    
    master_positive <- master_positive %>% 
      bind_rows(positive)
    
    message("Bottom pathway enrichment analyses for ", fav_gene)
    flat_bottom_complete <- dep_bottom %>%
      pull("gene") %>% 
      enrichr_loop(., focused_lib) %>%  #added small here
      top_n(100) 
    
    negative <- flat_bottom_complete %>% 
      mutate(fav_gene = fav_gene) %>% 
      group_by(fav_gene) %>% 
      nest()
    
    master_negative <- master_negative %>% 
      bind_rows(negative)

    #plot setup
    target_achilles <- achilles_long %>% 
      filter(gene == fav_gene) %>% 
      left_join(expression_join, by = "X1") %>% 
      select(stripped_cell_line_name, lineage, dep_score)
    
    target_achilles_top <- target_achilles %>% 
      top_frac(dep_score, n = 0.01)
    
    target_achilles_bottom <- target_achilles %>% 
      top_frac(dep_score, n = -0.01) %>% 
      arrange(dep_score)
    
    #plot1
    dep_plot1 <- ggplot(target_achilles) +
      geom_histogram(aes(x = dep_score), binwidth = 0.25, color = "lightgray") +
      labs(x = "Dependency Score (binned)") + 
      theme_light()
    
    #plot2
    dep_plot2 <- ggplot(target_achilles) +
      geom_point(aes(x = fct_rev(fct_reorder(target_achilles$stripped_cell_line_name,  target_achilles$dep_score, .desc = TRUE)), y = dep_score)) + #is target achilles in here?
      labs(x = "Cell Lines", y = "Dependency Score") +
      geom_hline(yintercept = mean_virtual_achilles) +
      geom_hline(yintercept = achilles_upper, linetype="dashed") +
      geom_hline(yintercept = achilles_lower, linetype="dashed") +
      geom_hline(yintercept = 0) +
      geom_point(data = target_achilles_top, aes(x = stripped_cell_line_name, y = dep_score), color = "red") +
      geom_point(data = target_achilles_bottom, aes(x = stripped_cell_line_name, y = dep_score), color = "red") +
      theme_light() +
      theme(axis.text.x=element_blank(), axis.ticks.x=element_blank()) + # axis.title.x=element_blank()
      NULL
    
    #place objects into list columns
    plot1 <- list(dep_plot1)
    plot2 <- list(dep_plot2)
    plots <- tibble(fav_gene, plot1, plot2)
    
    master_plots <- master_plots %>% 
      bind_rows(plots)
    
    } else {
    return()
  }
}

beep(sound = 8) 
#because mario is awesome
```

#save Rdata
```{r}
save(master_bottom_table, file=here::here("data", "master_bottom_table.RData"))
save(master_top_table, file=here::here("data", "master_top_table.RData"))
save(master_plots, file=here::here("data", "master_plots.RData"))
save(master_positive, file=here::here("data", "master_positive.RData"))
save(master_negative, file=here::here("data", "master_negative.RData"))
```

#load Rdata
```{r}
load(here::here("data", "achilles_cor.RData"))

master_bottom_table <- load(file=here::here("data", "master_bottom_table.RData"))
master_top_table <- load(file=here::here("data", "master_top_table.RData"))
master_plots <- load(file=here::here("data", "master_plots.RData"))
master_positive <- load(file=here::here("data", "master_positive.RData"))
master_negative <- load(file=here::here("data", "master_negative.RData"))
```

#test retreival to set objects
```{r}
#wrap this in if_else for if it's found v. not
fav_gene <- c("MFSD8")

fav_gene_summary <- gene_summary %>% 
      filter(approved_symbol == fav_gene)

dep_top <- master_top_table %>% 
  filter(fav_gene == !!fav_gene) %>% 
  unnest(data) %>% 
  select(-fav_gene) %>% 
  arrange(desc(r2))

dep_bottom <- master_bottom_table %>% 
  filter(fav_gene == !!fav_gene) %>% 
  unnest(data) %>% 
  select(-fav_gene) %>% 
  arrange(r2)

flat_top_complete <- master_postive %>% 
  filter(fav_gene == !!fav_gene) %>% 
  unnest(data) %>% 
  select(-fav_gene) %>% 
  arrange(Adjusted.P.value)

flat_bottom_complete <- master_negative %>% 
  filter(fav_gene == !!fav_gene) %>% 
  unnest(data) %>% 
  select(-fav_gene) %>% 
  arrange(Adjusted.P.value)

dep_plot1 <- master_plots %>% 
  filter(fav_gene == !!fav_gene) %>% 
  pluck("plot1")

dep_plot2 <- master_plots %>% 
  filter(fav_gene == !!fav_gene) %>% 
  pluck("plot2")

```

# test generation of report
```{r}
for (fav_gene in gene_group) {
  if(fav_gene %in% names(achilles_cor) == 1){ #this code checks to see if the gene is in the analysis, and if not, skips
    fav_gene_summary <- gene_summary %>% 
      filter(approved_symbol == fav_gene)
    
    dep_top <- master_top_table %>% 
      filter(fav_gene == !!fav_gene) %>% 
      unnest(data) %>% 
      select(-fav_gene) %>% 
      arrange(desc(r2))
    
    dep_bottom <- master_bottom_table %>% 
      filter(fav_gene == !!fav_gene) %>% 
      unnest(data) %>% 
      select(-fav_gene) %>% 
      arrange(r2)
    
    flat_top_complete <- master_postive %>% 
      filter(fav_gene == !!fav_gene) %>% 
      unnest(data) %>% 
      select(-fav_gene) %>% 
      arrange(Adjusted.P.value)
    
    flat_bottom_complete <- master_negative %>% 
      filter(fav_gene == !!fav_gene) %>% 
      unnest(data) %>% 
      select(-fav_gene) %>% 
      arrange(Adjusted.P.value)
    
    dep_plot1 <- master_plots %>% 
      filter(fav_gene == !!fav_gene) %>% 
      pluck("plot1")
    
    dep_plot2 <- master_plots %>% 
      filter(fav_gene == !!fav_gene) %>% 
      pluck("plot2")
    
    #render output
    render("report_depmap_complete.rmd", output_dir = here::here("results"), output_file = paste0(fav_gene, '_depmap.pdf'))
  } else {
    fav_gene_summary <- gene_summary %>% 
      filter(approved_symbol == fav_gene)
    
    #render output
    render("report_dummy_depmap.rmd", output_dir = here::here("results"), output_file = paste0(fav_gene, '_depmap.pdf'))
  }
}

#beep(sound = 8) 
#because mario is awesome
```

#how long?
```{r}
end_time <- Sys.time()
time_taken <- round(as.duration(start_time %--% end_time)/dminutes(1), digits = 1)
print(time_taken)
```
Approximate time to run was `r time_taken` minutes.

#print Session information for provenance and reproducibility
```{r}
utils:::print.sessionInfo(sessionInfo()[-8]) 
#You can remove an item from sessionInfo(), which is a list with a class attribute, by printing the resulting object omitting one of the list items (omitted list of packages installed, but not loaded)

```
#stamp
```{r}
stamp("Data updated December 31, 1979")(now())
```

#beep
```{r}
beep(sound = 8) #because mario is awesome
```



