---
output: html_fragment
---

```{r setup, include=FALSE}
###knit this document to generate methods.html, which then is served by the Shiny app; html fragments only contain <body></body> content, and are designed to be served in other pages###
##upon each release, run 
#generate_methods.R, 

#then knit
knitr::opts_chunk$set(echo = TRUE)
library(janitor)
library(lubridate)
library(tidyverse)
library(knitr)
library(cowplot)
library(viridis)
library(corrr)
library(here)
library(patchwork)
library(enrichR)
library(tidygraph)
library(ggraph)
library(patchwork)
library(ggrepel)
library(kableExtra)
library(visNetwork)
library(htmltools)
```

```{r message=FALSE, warning=FALSE, include=FALSE}
#clear environment
#rm(list=ls()) 

time_begin_data <- Sys.time()

#read current release information 
source(here::here("code", "current_release.R"))

#Load functions
source(here::here("code", "fix_names.R"))

##LOAD DATA
fav_gene <- "TP53"
gene_summary <- readRDS(file = here::here("data", paste0(release, "_gene_summary.Rds")))
achilles <- readRDS(file = here::here("data", paste0(release, "_achilles.Rds")))
achilles_cor <- readRDS(file = here::here("data", paste0(release, "_achilles_cor.Rds")))
achilles_cor_long <- achilles_cor %>% 
  pivot_longer(-rowname, names_to = "achilles_gene", values_to = "achilles_correlation")
expression <- readRDS(file = here::here("data", paste0(release, "_expression.Rds")))
expression_join <- readRDS(file = here::here("data", paste0(release, "_expression_names.Rds"))) #fix the code below...can I make it use the functions, instead of manually calling them here?
pubmed_concept_pairs <- readRDS(file = here::here("data", paste0(release, "_pubmed_concept_pairs.Rds")))

#read data from generate_depmap_pathways.R
master_bottom_table <- readRDS(file=here::here("data", paste0(release, "_master_bottom_table.Rds")))
master_top_table <- readRDS(file=here::here("data", paste0(release, "_master_top_table.Rds")))

master_positive <- readRDS(file=here::here("data", paste0(release, "_master_positive.Rds")))
master_negative <- readRDS(file=here::here("data", paste0(release, "_master_negative.Rds")))

#LOAD STATS
sd_threshold <- readRDS(file = here::here("data", paste0(release, "_sd_threshold.rds")))
achilles_lower <- readRDS(file = here::here("data", paste0(release, "_achilles_lower.rds")))
achilles_upper <- readRDS(file = here::here("data", paste0(release, "_achilles_upper.rds")))
mean_virtual_achilles <- readRDS(file = here::here("data", paste0(release, "_mean_virtual_achilles.rds")))
sd_virtual_achilles <- readRDS(file = here::here("data", paste0(release, "_sd_virtual_achilles.rds")))
mean_virtual_expression <- 0 #set to zero for now, to prevent knitting error; fix when you add this to the methods
expression_upper <- 0 #set to zero for now, to prevent knitting error; fix when you add this to the methods

#generate raw data too
achilles_raw <- read_csv(achilles_url, col_names = TRUE) %>% 
  `colnames<-`(str_remove_all(names(.), "\\s\\(\\d+\\)")) %>% 
  rename(X1 = 1)

achilles_raw <- clean_colnames(achilles_raw)


source(here::here("code", "fun_tables.R"))
source(here::here("code", "fun_plots.R"))
source(here::here("code", "fun_graphs.R"))
```

![](`r "https://source.unsplash.com/y41FEMqdJ3A/800x600"`)  
  
## Why this project?
Identifying the key functions of human genes is a major biomedical research goal. While some genes are very well-studied, most human genes we know little about. Mapping a broad set of gene functions *en masse* is currently not possible, and therefore most scientific research proceeds with a one-gene-at-a-time approach.  While this reductionist approach has guided the scientific methods for hundreds of years, the volume, complexity, and sophistication of modern science necessitate alternative approaches.  
  
  
```{r pubmed, echo=FALSE, message=FALSE, warning=FALSE}
gene2pubmed <- read_tsv(gene2pubmed_url, col_names = TRUE) %>% #gene2pubmed_url read from current_release.R
  clean_names() %>%
  filter(number_tax_id == 9606) %>%  #only the rows corresponding to humans (#tax_id = 9606) 
  group_by(gene_id) %>% 
  count(sort = TRUE) %>% 
  left_join(gene_summary, by = c("gene_id" = "ncbi_gene_id")) %>% 
  select(gene_id, approved_symbol, n, approved_name) %>% 
  filter(!is.na(approved_symbol))

ggplot() +
  geom_point(data = gene2pubmed, mapping = aes(x = fct_reorder(approved_symbol, n), y = n), alpha = 0.2) +
  geom_text_repel(data = subset(gene2pubmed, n > 3000), mapping = aes(x = fct_reorder(approved_symbol, n), y = n, label = approved_symbol)) +
  labs(x = "Human Gene", y = "Number of Publications", subtitle = "Number of Publications per human gene") +
  theme_cowplot() +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  coord_cartesian(xlim = c(0, 33000))
```
  
  
New tools in Data Science -- a combination of computer programming, math & statistics, and topical expertise -- combined with the rapid adoption of open science and data sharing allow scientists to access publicly available datasets and interrogate these data *before* performing any experiments.  
  
Such tools allow researchers to identify new hypotheses for well-studied genes, or new processes for un-annotated genes. Indeed, most human genes are poorly studied, perhaps because of the intractability to study them with no prior knowledge.  
  
This project aims to address this challenge. Welcome to data-driven hypothesis.  
   
## What is this project?
The overall goal of the data-driven hypothesis (DDH) project is to use new tools in Data Science to generate hypotheses supported by data that can be tested in the lab. Several high-quality, functional genomic datasets are published online and made available with Creative Commons Attribution 4.0 International [(CC BY 4.0)](https://creativecommons.org/licenses/by/4.0/) licenses. Functional genomics is a field of molecular biology that aims to understand the function of all genes and proteins in a genome -- a stated goal of much basic science research. In functional genomics, experimental strategies generally involve high-throughput, genome-wide approaches rather than a more traditional “gene-by-gene” approach. The advent and rapid adoption of data-sharing platforms, such as [figshare.com](https://figshare.com) have provided high-quality data sets for public interrogation.  
  
We present here a comprehensive resource for functionally mapping thousands of genes across the human genome. Importantly, this method integrates gene essentiality, gene expression, and literature mining to identify candidate functions of known or unknown genes.  The DDH project holds tremendous promise to generate hypotheses, data, and knowledge in order to provide a deep understanding of the dynamic properties of an organism.  We provide this resource via an intuitive online interface, which will provide the scientific community a platform to query and prioritize experimental hypotheses to test in the lab.  
  
## How does it work?
This project began as an extension of a simple and common concept in molecular biology called gene co-expression analysis. When a gene of unknown function is identified, one strategy to learn something about the new gene is to identify shared patterns of expression with other genes. If unknown Gene X is expressed with known genes A, B, and C, then you can infer that Gene X might be part of a functional module with A, B, C. This approach is particularly powerful when genes A, B, and C are part of a known biological pathway, which leads to the hypothesis that Gene X might also be part of that same pathway.  
  
```{r gene_coexpression, echo=FALSE, message=FALSE, warning=FALSE}
# Dummy data
x <- paste0("CTL ", seq(1,3))
x <- x %>% 
  c(x <- paste0("EXP ", seq(1,3)))
y <- paste0("gene ", LETTERS[24])
y <- y %>% 
  c(paste0("gene ", LETTERS[1:6]))
data <- expand.grid(X=x, Y=y, stringsAsFactors = TRUE)

pos <- runif(100, 0.2, 1)
neg <- runif(100, -1, -0.2)

i <- 0
A <- NULL
while (i < 4) {
  a <- sample(pos, 3)
  a <- a %>% c(sample(neg,3))
  A <- c(A, a)
  i <- i + 1
}
j <- 0
B <- NULL
while (j < 3) {
  b <- sample(neg, 3)
  b <- b %>% c(sample(pos,3))
  B <- c(B, b)
  j <- j + 1
}

data$Z <- c(A,B)

# Heatmap 
heatmap_gene <- ggplot(data, aes(X, fct_rev(Y), fill= Z)) + 
  geom_tile() +
  labs(x = "", y = "", fill = "Expression \nCorrelation", subtitle = "Gene Coexpression Example*", caption = "*simulated data") +
  theme_minimal_grid(color = "white") +
  scale_x_discrete(position = "top")

heatmap_gene
ggsave(here::here("results", "fig1a_heatmap_gene.png"), plot = heatmap_gene, dpi = 300)
```
  
Following on this idea, we set out to map genes to common functional pathways based on dependence of a pathway for cell viability. Project Achilles is a systematic effort by the [Broad Institute](https://www.broadinstitute.org) as part of a larger ['DepMap' project](http://www.depmap.org) aimed at identifying and cataloging gene essentiality across hundreds of well-characterized cancer cell lines using highly standardized pooled genome-scale loss-of-function screens. This project uses lentiviral-based pooled RNAi or CRISPR/Cas9 libraries to systematically knock-out each gene in the genome, which allows for the stable suppression/ablation of each gene individually in a subset of cells within a pooled format allowing for genome wide interrogation of gene essentiality. Using computational modeling, a normalized value of gene essentiality is given for each gene in a single cell line. A lower score means that a gene is more likely to be essential in a given cell line. A score of -1 corresponds to the median of all common essential genes, whereas a score of 0 is equivalent to a gene that is not essential; a positive score indicates a gain in fitness and often identifies tumor suppressor genes.  
  
It is well-known that human cancer cell lines rely on different pathways for their viability. Indeed this is the entire rationale for personalized, precision medicine in cancer. The overall goal of the 'DepMap' project is to identify all essential genes in 2000 cell lines over the 5-year project period to identify new therapeutic targets in various cancers. Despite not knowing the mechanistic basis for why some cell lines require specific genes while other cell lines do not, we reasoned that intrinsic reliance of a cell on a pathway might allow unbiased detection of novel genes participating in specific pathways.  
  
```{r cell_dependency, echo=FALSE, message=FALSE, warning=FALSE}
# Dummy data
x <- paste0("cell line ", seq(1,6))
y <- paste0("gene ", LETTERS[24])
y <- y %>% 
  c(paste0("gene ", LETTERS[1:6]))
data <- expand.grid(X=x, Y=y, stringsAsFactors = TRUE)

pos <- runif(100, 0.2, 1)
neg <- runif(100, -1, -0.2)

i <- 0
A <- NULL
while (i < 3) {
  a <- sample(pos, 3)
  a <- a %>% c(sample(neg,3))
  A <- c(A, a)
  i <- i + 1
}
j <- 0
B <- NULL
while (j < 4) {
  b <- sample(neg, 3)
  b <- b %>% c(sample(pos,3))
  B <- c(B, b)
  j <- j + 1
}

data$Z <- c(B,A)

# Heatmap 
heatmap_dep <- ggplot(data, aes(X, fct_rev(Y), fill= Z)) + 
  geom_tile() +
  labs(x = "", y = "", fill = "Dependency \nCorrelation", subtitle = "Dependency Score Example*", caption = "*simulated data") +
  theme_minimal_grid(color = "white") + 
  scale_fill_viridis() + 
  scale_x_discrete(position = "top")

heatmap_dep
ggsave(here::here("results", "fig1b_heatmap_dep.png"), plot = heatmap_dep, dpi = 300)
```
  
## What did I do?
Essential gene data from Project Achilles were downloaded from the DepMap portal at:  [depmap.org](https://depmap.org/portal/download/). The `r release` release contains gene essentiality scores for `r ncol(achilles)` genes across `r nrow(achilles)` cell lines, and was used for this project. 
  
```{r dep_scores, echo=FALSE, message=FALSE, warning=FALSE}
#insert summary of scores
achilles_raw_long <- achilles_raw %>% 
  pivot_longer(-X1, names_to = "gene", values_to = "dep_score") %>% 
  filter(!is.na(dep_score))

depmap_raw <- ggplot() +
  geom_histogram(data = achilles_raw_long, aes(x = dep_score), color = "gray", fill = "#02224C", binwidth = 0.1) +
  labs(x = "Dependency Score", y = "Count (Million)", subtitle = paste0("Dependency Score Distribution from DepMap release ", release)) +
  scale_y_continuous(breaks = c(0, 500000, 1000000, 1500000, 2000000), labels = c("0", "0.5", "1", "1.5", "2")) +
  scale_x_continuous(breaks = c(-5:7)) +
  theme_cowplot()

depmap_raw
ggsave(here::here("results", "fig2a_depmap_raw.png"), plot = depmap_raw, dpi = 300)
```
  
#### Predictive analytics by pattern recognition
To find patterns in gene dependencies across cell lines, we generated a Pearson correlation matrix of all genes by all genes. This analysis generated gene-gene correlation values that matched values published on [depmap.org](https://depmap.org), validating the first step in our analysis. High levels of gene expression are often thought to be indicative of key genes for a given cell type. Thus, we next compared dependency values to gene expression values. The [Cancer Cell Line Encyclopedia](https://portals.broadinstitute.org/ccle/about) project is a collaboration between the Broad Institute, and the Novartis Institutes for Biomedical Research and its Genomics Institute of the Novartis Research Foundation, which together conduct detailed genetic and pharmacologic characterization of a large panel of human cancer models. As of the CCLE 2019 release, `r nrow(expression)` cell lines have been characterized for gene expression. In the `r release` DepMap release, `r sum(expression$X1 %in% achilles$X1 == TRUE)` of the `r nrow(achilles)` cell lines have gene expression data. Using these two datasets, we compared the essentiality of a gene to its expression value.  
  
```{r depVexp, echo=FALSE, message=FALSE, warning=FALSE}
expression_long <- expression %>% 
  filter(expression$X1 %in% achilles$X1 == TRUE) %>% #matches cells and makes sure I'm analysing CCLE cells in Achilles
  pivot_longer(-X1, names_to = "gene", values_to = "gene_expression") %>% 
  arrange(desc(gene_expression))

joined <- expression_long %>% 
  left_join(achilles_raw_long, by = c("gene" = "gene", "X1" = "X1"))

log_breaks <- c(1, 10, 100, 1000, 10000, 100000, 1000000)
log_labels <-c("1", "10", "100", "1,000", "10,000", "100,000", "1,000,000")
depVexp <- joined %>% 
  ggplot() +
  geom_hex(aes(x = dep_score, y = gene_expression), 
           bins = 90, na.rm = TRUE, size = 0.2, color = "black") +
  scale_fill_viridis_c(option = "E", name = "Hex Bin \nCount", trans = "log10", breaks = log_breaks, labels = log_labels, limits = c(1,1000000)) + #log scale transform legend
  labs(x = "Dependency Score", y = "Gene Expression", subtitle = "Gene expression is unrelated to its essentiality") +  
  theme_cowplot()

depVexp
ggsave(here::here("results", "fig2b_depVexp.png"), plot = depVexp, dpi = 300)
```
  
We predicted a V-shaped curve, with stronger dependencies as gene expression increases. Surprisingly, we saw no relationship between gene expression and gene essentiality, where genes with both low and high expression displayed both gains and losses in fitness. The overall observation from this dataset shows baseline gene expression levels are poor indicators of the essentiality of a gene. This analysis also highlighted that several genes were binned on the x-axis, i.e. could have no measurable expression levels, but have assigned dependency scores. Across `r nrow(achilles)` cell lines in the Achilles project, `r round(((expression_long %>% filter(gene_expression == 0) %>% count())/(length(expression_long$gene_expression))*100), digits = 1)`% of all gene expression values are zero, confirming this notion. 
  
#### Noise Reduction
Given cells do not express all genes, but might receive a dependency score in this experimental paradigm, we sought to remove dependency scores for gene-cell line pairs that have an expression value of zero under basal conditions. Of the `r nrow(achilles)` cell lines for which gene essentiality data is collected, `r sum(achilles$X1 %in% expression$X1)` have genome-wide gene expression data. From these cell lines, we removed dependency scores for genes from cell line that have a corresponding gene expression value of zero.  
  
```{r expression_0, echo=FALSE, message=FALSE, warning=FALSE}
#how many 0s?
expression_0 <- ggplot() +
  geom_histogram(data = expression_long, aes(x = gene_expression), color = "gray", fill = "#02224C", binwidth = 0.25) +
  labs(title = paste("Binned gene expression values in ", sum(achilles$X1 %in% expression$X1), " Achilles cell lines", sep = ""), 
       subtitle = paste0(
    round(((expression_long %>% filter(gene_expression == 0) %>% count())/(length(expression_long$gene_expression))*100), digits = 1), "% of all values are zero"), 
    x = "Gene expression (RPKM)", 
    y = "Count (Million)") +
  scale_y_continuous(breaks = seq(0, 3000000, by = 1000000), labels = c("0", "1", "2", "3")) +
  theme_cowplot() +
  NULL

expression_0
ggsave(here::here("results", "fig2c_expression_0.png"), plot = expression_0, dpi = 300)
```
  
For some genes expressed in highly specific and restricted cell types, this operation removed many dependency values. After removing these values, we found that highly specialized genes in discrete cell types have too few cells with both gene expression values and gene essentiality values to assign a meaningful correlation value. Thus, if a gene was absent from too many cell lines, we omitted it to prevent assigned values from relying on too few data points.   
  
```{r dep_0s, echo=FALSE, message=FALSE, warning=FALSE}
#gets rid of zeros by replacing with NAs
no_expression <- expression_long %>% 
  filter(gene_expression == 0) %>% 
  unite(X1, gene, col = "match", sep = "-", remove = TRUE) %>% 
  pull(match)

achilles_no0 <- achilles_raw_long %>% 
  unite(X1, gene, col = "match", sep = "-", remove = FALSE) %>% 
  filter(match %in% no_expression == FALSE) %>% 
  select(-match) %>%
  spread(gene, dep_score)

#count NAs in corrected dataframe
achilles_no0_plot <- achilles_no0 %>% 
  summarise_all(list(~sum(is.na(.)))) %>% 
  gather(gene, NAs) %>% 
  arrange(desc(NAs)) %>% 
  mutate(pos = sum(achilles$X1 %in% expression$X1)-NAs)

#this is code to draw arrows graph labels
arrows <- tibble(
  x1 = c(sum(achilles$X1 %in% expression$X1)-na_cutoff, sum(achilles$X1 %in% expression$X1)-na_cutoff),
  x2 = c(375, 50),
  y1 = c(6000, 8000),
  y2 = c(2500, 5000)
)

threshold <- ggplot(achilles_no0_plot) +
  geom_histogram(aes(x = pos), color = "gray", fill = "#02224C") +
  labs(x = "# of cell lines expressing target gene", y = "# of genes in bin", subtitle = "The number of cell lines each gene was present within") +
  geom_curve(data = arrows, aes(x = x1, y = y1, xend = x2, yend = y2),
             arrow = arrow(length = unit(0.07, "inch")), size = 0.4,
             color = "gray60", curvature = .20) +
  annotate("text", x = 35, y = 3000, size = 4, color = "gray60",
           label = glue::glue("These genes \nhave too few \ncells with \nexpression")) +
  annotate("text", x = 500, y = 2500, size = 4, color = "gray60",
           label = glue::glue("These genes have several \ncell lines with expression")) +
  annotate("text", x = sum(achilles$X1 %in% expression$X1)-na_cutoff + 5, y = 12000, size = 3, color = "red", hjust = 0, fontface = "bold",
           label = glue::glue("Threshold")) +
  geom_vline(aes(xintercept = sum(achilles$X1 %in% expression$X1)-na_cutoff), color = "red") +
  theme_cowplot()

#code below sets threshold to remove too many NAs from raw data without zeros
toomanyNAs <- achilles_no0 %>% 
  summarise_all(list(~sum(is.na(.)))) %>% 
  gather(gene, NAs) %>% 
  arrange(desc(NAs)) %>% 
  filter(NAs > na_cutoff) %>% 
  pull(gene)

achilles_clean <- achilles_no0 %>% select(-one_of(toomanyNAs)) #this is now the same object as achilles, but different than achilles_raw; referenced below
#would be nice to list these omitted genes, but lose the kable in Rmd->md->html conversion

##relationship between number of cell lines that drive gene correlation and number of genes in list
count_v_max_no0 <- achilles_no0 %>% 
  select(-X1) %>% 
  correlate() %>% 
  pivot_longer(-1, names_to = "genes_no0", values_to = "correlations_no0") %>% 
  rename(rowname = 1) %>% 
  filter(correlations_no0 > achilles_upper) %>%  #| achilles_correlation_raw < achilles_lower) %>% 
  dplyr::group_by(rowname) %>% 
  summarize(count = n(), 
            max = max(correlations_no0)) %>% 
  left_join(achilles_no0_plot, by = c("rowname" = "gene")) %>% 
  arrange(pos)

sampling <- count_v_max_no0 %>% 
  ggplot() +
  geom_point(aes(x = pos, y = count), alpha = 0.2, color = "#02224C") +
  geom_vline(xintercept = nrow(achilles_clean) - na_cutoff, color = "red") +
  #geom_vline(xintercept = nrow(achilles_clean) - na_cutoff_tmp, color = "red", linetype="dashed") +
  labs(y = "# of genes correlated w/ target", x = "# of cell lines expressing target gene", subtitle = "Sampling too few cells weakens predictions") +
  annotate("text", x = sum(achilles$X1 %in% expression$X1)-na_cutoff + 10, y = 10, size = 3, color = "red", hjust = 0, fontface = "bold",
           label = glue::glue("Threshold")) +
  theme_cowplot()

sampling
ggsave(here::here("results", "fig2e_sampling.png"), plot = sampling, dpi = 300)
```
  
  
We set a threshold of no more than `r na_cutoff` zeros, meaning that if a gene had fewer than `r nrow(achilles)-na_cutoff` cell lines with dependency values, the correlation pattern of a gene would be meaningless, and that gene was therefore removed. This value was based on removing `r paste0(round(fraction_cutoff*100, 0), " %")` of the genes with the fewest number of cells deriving the correlation value and having the highest number of correlated gene patterns. This process removed `r length(toomanyNAs)`  genes that had too few cells with expression and dependency data.  
  
```{r raw, echo=FALSE, message=FALSE, warning=FALSE}
threshold
ggsave(here::here("results", "fig2e_threshold.png"), plot = threshold, dpi = 300)

```
  
To quantify the effect of these cleaning step on the data, we compared the correlation matrix generated from the raw data to the correlation matrix from the cleaned data. Plotting these ~200M gene-gene pairs revealed that more positive genes became more positive compared positive genes that became negative; conversely, more gene pairs with negative correlations became more negative rather than less negative. Simply stated, these cleaning steps had a greater effect on strengthening the data in either direction rather than weakening it, supporting these steps as important for this resource. These 'cleaned' dependency data had `r length(achilles_clean)` remaining gene-dependency pairs, which were then used to generate correlation matrix.  
  
```{r special_sauce, echo=FALSE, message=FALSE, warning=FALSE}
#generate cor matrix from raw data to compare raw correlation to final correlation
achilles_cor_raw <- achilles_raw %>% 
  select(-X1) %>% 
  correlate() 

#shave to reduce redundancies
achilles_cor_shave <- shave(achilles_cor)
achilles_cor_raw_shave <- shave(achilles_cor_raw)

#pivot both longer, but use shaved dataset to plot only one gene-gene pair
achilles_cor_shave_long <- achilles_cor_shave %>% 
  pivot_longer(-1, names_to = "achilles_gene", values_to = "achilles_correlation") %>% 
  filter(!is.na(achilles_correlation))

achilles_cor_raw_shave_long <- achilles_cor_raw_shave %>% 
  rename(rowname = 1) %>% 
  pivot_longer(-rowname, names_to = "achilles_gene", values_to = "achilles_correlation_raw") %>% 
  filter(!is.na(achilles_correlation_raw))

achilles_joined <- achilles_cor_raw_shave_long %>% 
  left_join((achilles_cor_shave_long), by = c("rowname" = "rowname", "achilles_gene" = "achilles_gene"))

log_breaks10M <- c(1, 10, 100, 1000, 10000, 100000, 1000000, 10000000)
log_labels10M <-c("1", "10", "100", "1,000", "10,000", "100,000", "1,000,000", "10,000,000")

effect_of_cleaning <- achilles_joined %>% 
  #sample_n(1000000) %>% 
  ggplot() +
  geom_abline(intercept = 0, slope = 1, color = "gray") +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_hex(aes(x = achilles_correlation_raw, y = achilles_correlation), 
           bins = 90, na.rm = TRUE, size = 0.2, color = "black") +
  scale_fill_viridis_c(option = "E", name = "Hex Bin \nCount", trans = "log10", breaks = log_breaks10M, labels = log_labels10M, limits = c(1,10000000)) + #breaks and labels defined above
  labs(x = "Raw Correlation", y = "Corrected Correlation", subtitle = "Effect of cleaning") +  
  coord_fixed(ratio = 1, xlim = c(-1, 1), ylim = c(-1, 1)) +
  theme_minimal()

#achilles_cor read in line 35
#achilles_cor_long generated line 36

#achilles_cor_heatmap <- achilles_cor %>% 
#  #rearrange(absolute = FALSE) %>%    
#  pivot_longer(-rowname, names_to = "achilles_gene", values_to = "achilles_correlation") %>% 
#  filter(!is.na(achilles_correlation)) %>% 
#  ggplot(aes(rowname, achilles_gene, fill = achilles_correlation)) +
#  geom_raster() +
#  scale_fill_viridis(option = "D", name="Pearson\nCorrelation", limit = c(-1, 1)) +
#  labs(title = "Gene-gene correlation matrix") +
#  theme_void() +
#  coord_fixed()

effect_of_cleaning #+ achilles_cor_heatmap
ggsave(here::here("results", "fig2f_effect_of_cleaning.png"), plot = effect_of_cleaning, dpi = 300)

```

  
## How does it work?
To identify genes that shared similar patterns of essentiality with other genes, thereby placing genes in functional pathways, we generated a gene-by-gene Pearson correlation matrix on these prioritized data to quantify the similarity in dependency patterns and annotate genes in functional pathways.  

```{r r2, echo=FALSE, message=FALSE, warning=FALSE}
r2 <- achilles_cor_long %>% 
  filter(!is.na(achilles_correlation)) %>% 
  #sample_n(1000) %>% 
  ggplot() +
  geom_histogram(aes(x = achilles_correlation), color = "gray", fill = "#02224C", binwidth = 0.05) +
  scale_y_continuous(breaks = c(0, 25000000, 50000000, 75000000), labels = c("0", "25", "50", "75")) +
  labs(x = "r^2 values", y = "Count (Million)", subtitle = "Distribution of correlation values across all gene-gene pairs") +
  theme_cowplot()

r2
ggsave(here::here("results", "fig3a_r2.png"), plot = r2, dpi = 300)
```
  
This process generated approximately `r round(nrow(achilles_cor_long)/1000000, 0)` million correlation values, with a distribution centered around zero. This output produced a range of maximum correlation values for each gene.  
  
```{r achilles_max, echo=FALSE, message=FALSE, warning=FALSE}
achilles_max <- achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) %>% 
  ggplot(aes(max)) +
  geom_histogram(binwidth = 0.05, color = "gray", fill = "#02224C") + 
  #geom_vline(aes(xintercept = achilles_upper), color = "red") +
  labs(subtitle = "Distribution of maximum r^2 values for each gene", x = "Maximum r^2 values", y = "Count") +
  theme_cowplot()

achilles_max
ggsave(here::here("results", "fig3b_achilles_max.png"), plot = achilles_max, dpi = 300)
```
  
#### Statistics
Rather than setting an arbitrary threshold for the r^2 value that would be considered a low, medium, or high correlation between two genes, we performed a permutation test on the correlated data. A permutation test involves permuting one or more variables in a data set before performing the test, in order to break any existing relationships and simulate the null hypothesis. In this case, we broke the relationship between gene-gene pairs and the correlation values. We then generated a distribution of null statistics (fake means), along with standard deviations of these sampled data. This strategy identified at which correlation value to draw a threshold of a "significant correlation" for these analyses. We sampled 20,000 r^2 values from all gene-gene pairs without replacement simulating a virtual Achilles correlation dataset for a single cell. We then repeated this process 1000 times mimicking 1000 discrete cell lines.  
  
This statistical analysis produced the following data:  
**Mean:** `r mean_virtual_achilles`  
**Standard Deviation:** `r sd_virtual_achilles`  
  
Using a standard deviation threshold of `r sd_threshold`, we calculated the boundaries of r^2 values to be greater than `r round(achilles_upper, 2)` or lower than `r round(achilles_lower, 2)` for negative correlations. Simply stated, gene-gene correlations above or below, respectively, these values were considered strong and used for subsequent analyses. The r^2 values and corresponding z-score transformations of these values are shown on the tables.  
  
#### Comparing to Known Literature    
The process of generating a prioritized gene list based on similar patterns to a query gene is an unbiased approach to map genes into functional pathways. To find novel genes in established pathways, this approach requires identifying "novel" or "unexpected" genes in the gene list. However, what may seem unexpected to one researcher might be obvious or intuitive to another. To remove bias in prioritizing candidate genes for further study and to programatically identify novelty, we generated a **co-publication index**, where every human gene was compared to publication co-incidence with every other human gene in the [PubMed](https://pubmed.ncbi.nlm.nih.gov) database.  
  
PubMed contains 30 million publication records containing article metadata and abstracts. We began by first parsing each of the 30 million abstracts for all gene names and gene aliases in an abstract. Next we were able to generate article level summaries for each gene in each article, and then counted gene-gene co-occurrence in an abstract. Measuring all human gene-gene co-occurrences across all PubMed records allowed us to identify the frequency with which a gene was published with another gene. Unfortunately, gene nomenclature is a mess, with official gene names changing over time, and several genes sharing common aliases. As a keystone example, the protein p38 has been used as a name to describe the genes MAPK14 (NCBI gene id 1432), AHSA1 (NCBI gene id 10598), and AIMP2 (NCBI gene id 7965). The protein p38 associated with MAPK14 is among the most studied and published on genes in the human genome, whereas AHSA1 and AIMP2 are among the least studied. This discrepancy and ambiguity in identifying genes in abstracts necessitated an alternative approach.  
  
Next, we leveraged a PubMed resource from the National Library of Medicine (NLM) called gene2pubmed that links unique NCBI gene ids to a PubMed id. This resource is part of the NLM's Indexing Initiative (IND) is working to automate indexing methods partially or completely manual indexing practices. Using gene2pubmed, we again measured all human gene-gene co-occurrences across all PubMed records. While this method overcomes the challenges associated with gene nomenclature by providing gene ids, it is still limited in its ability to curate genes associated with scientific papers by the same challenges in gene nomenclature.  
  
A new machine learning (ML)-based resource developed by the Computational Biology Branch at the NLM/NCBI called [Pubtator](https://www.ncbi.nlm.nih.gov/research/pubtator/) aims to overcome these challenges by providing automatic annotations of biomedical concepts such as genes and mutations in PubMed abstracts and PMC full-text articles. Using advanced text-mining techniques, Pubtator goes beyond assigning genes to paper by developing a ML model that generates automated annotations for genes/proteins, genetic variants, diseases, chemicals, species and cell lines in a single paper, across all PubMed (30 million) abstracts and the PMC Text Mining subset of (3 million full text) articles.Improved concept identification systems and a new disambiguation module based on deep learning provides best-in-class annotation accuracy. Using these data, we filtered "concept ids" that matched official NCBI gene IDs, and then took the same approach described above to determine the number of times a gene was annotated with scientific paper, and the pairwise count of each gene with each other gene in a single paper, across all papers. Because of the wide range of annotated gene co-occurrences (`r paste0("0 to ", pubmed_concept_pairs %>% filter(target_gene == "INS") %>% unnest() %>% summarize(max(n)) %>% pull())`), we also generated a relative index [(co-occurrence count/max co-occurrence)*100] normalized to 100. The co-occurrence count and co-occurrence index provides useful, and unbiased information about the level of prior knowledge for gene-gene relationship identified by the analytic approach detailed above.  
  
#### Pathway Analyses
```{r focused_lib, echo=FALSE, message=FALSE, warning=FALSE}
focused_lib <- c("Achilles_fitness_decrease", "Achilles_fitness_increase", "Aging_Perturbations_from_GEO_down", "Aging_Perturbations_from_GEO_up", "Allen_Brain_Atlas_down", "Allen_Brain_Atlas_up", "ARCHS4_Cell-lines", "ARCHS4_IDG_Coexp", "ARCHS4_Kinases_Coexp", "ARCHS4_TFs_Coexp", "ARCHS4_Tissues", "BioCarta_2016", "BioPlex_2017", "Cancer_Cell_Line_Encyclopedia", "ChEA_2016", "Chromosome_Location_hg19", "CORUM", "Data_Acquisition_Method_Most_Popular_Genes", "Disease_Perturbations_from_GEO_down", "Disease_Perturbations_from_GEO_up", "Disease_Signatures_from_GEO_up_2014", "Drug_Perturbations_from_GEO_down", "Drug_Perturbations_from_GEO_up", "DrugMatrix", "DSigDB", "ENCODE_and_ChEA_Consensus_TFs_from_ChIP-X", "ENCODE_Histone_Modifications_2015", "ENCODE_TF_ChIP-seq_2015", "Enrichr_Libraries_Most_Popular_Genes", "Enrichr_Submissions_TF-Gene_Coocurrence", "Epigenomics_Roadmap_HM_ChIP-seq", "ESCAPE", "GeneSigDB", "GO_Biological_Process_2018", "GO_Cellular_Component_2018", "GO_Molecular_Function_2018", "GTEx_Tissue_Sample_Gene_Expression_Profiles_down", "GTEx_Tissue_Sample_Gene_Expression_Profiles_up", "GWAS_Catalog_2019", "HMDB_Metabolites", "HomoloGene", "Human_Gene_Atlas", "Human_Phenotype_Ontology", "HumanCyc_2015", "HumanCyc_2016", "huMAP", "InterPro_Domains_2019", "Jensen_COMPARTMENTS", "Jensen_DISEASES", "Jensen_TISSUES", "KEA_2015", "KEGG_2019_Human", "KEGG_2019_Mouse", "Kinase_Perturbations_from_GEO_down", "Kinase_Perturbations_from_GEO_up", "Ligand_Perturbations_from_GEO_down", "Ligand_Perturbations_from_GEO_up", "LINCS_L1000_Chem_Pert_down", "LINCS_L1000_Chem_Pert_up", "LINCS_L1000_Kinase_Perturbations_down", "LINCS_L1000_Kinase_Perturbations_up", "LINCS_L1000_Ligand_Perturbations_down", "LINCS_L1000_Ligand_Perturbations_up", "MCF7_Perturbations_from_GEO_down", "MCF7_Perturbations_from_GEO_up", "MGI_Mammalian_Phenotype_Level_4_2019", "Microbe_Perturbations_from_GEO_down", "Microbe_Perturbations_from_GEO_up", "miRTarBase_2017", "Mouse_Gene_Atlas", "MSigDB_Computational", "MSigDB_Oncogenic_Signatures", "NCI-60_Cancer_Cell_Lines", "NURSA_Human_Endogenous_Complexome", "Old_CMAP_down", "Old_CMAP_up", "OMIM_Disease", "OMIM_Expanded", "Panther_2016", "Pfam_Domains_2019", "Pfam_InterPro_Domains", "Phosphatase_Substrates_from_DEPOD", "PPI_Hub_Proteins", "Rare_Diseases_AutoRIF_ARCHS4_Predictions", "Rare_Diseases_AutoRIF_Gene_Lists", "Rare_Diseases_GeneRIF_ARCHS4_Predictions", "Rare_Diseases_GeneRIF_Gene_Lists", "Reactome_2016", "RNA-Seq_Disease_Gene_and_Drug_Signatures_from_GEO", "SILAC_Phosphoproteomics", "Single_Gene_Perturbations_from_GEO_down", "Single_Gene_Perturbations_from_GEO_up", "SubCell_BarCode", "SysMyo_Muscle_Gene_Sets", "TargetScan_microRNA_2017", "TF_Perturbations_Followed_by_Expression", "TF-LOF_Expression_from_GEO", "Tissue_Protein_Expression_from_Human_Proteome_Map", "Tissue_Protein_Expression_from_ProteomicsDB", "Transcription_Factor_PPIs", "TRANSFAC_and_JASPAR_PWMs", "TRRUST_Transcription_Factors_2019", "UK_Biobank_GWAS", "Virus_Perturbations_from_GEO_down", "Virus_Perturbations_from_GEO_up", "VirusMINT", "WikiPathways_2019_Human", "WikiPathways_2019_Mouse")
#find a way to out put this in a simple/clean format?
```
To identify clusters of genes with shared relationships, we performed gene set enrichment analysis. Enrichment analysis is a computational method for inferring knowledge about a target gene set by comparing it to annotated gene sets representing prior biological knowledge. Enrichment analysis determines whether an input set of genes significantly overlaps with annotated gene sets. For each gene in our matrix, we determined the number of genes that were greater than or less than `r sd_threshold` standard deviations away from the permuted mean. This target gene list was then queried against `r length(focused_lib)` gene sets across a [broad range of curated data](https://amp.pharm.mssm.edu/Enrichr/#stats). By leveraging the [Enrichr](https://amp.pharm.mssm.edu/Enrichr/) resource from the [Ma'ayan Laboratory](http://labs.icahn.mssm.edu/maayanlab/), we determined the top ranked pathways, processes, drugs, cell lines, tissues, or diseases, and ranked by p-value. In this setting, the p-value is computed using a standard statistical method used by most enrichment analysis tools: Fisher's exact test or the hypergeometric test. This is a binomial proportion test that assumes a binomial distribution and independence for probability of any gene belonging to any set. [See here for more information about how Enrichr computes its associations](https://amp.pharm.mssm.edu/Enrichr/help#background).  
    
#### Network Analyses
As an alternative approach to prioritize further study of novel functional mappings, an interactive network graph of gene-gene relationships can be built. A single gene query results in a list of top and bottom associated genes based on patterns of gene dependency scores. The top 10 and bottom 10 genes are then queried for each of their top and bottom correlated genes. These resulting 400 gene pairs (20 genes * 20 genes) are then used to build a network graph, which can reveal if a top associated gene with the query gene, also has the query gene in its top (or bottom) 10. In this way, strong functional networks can be identified and prioritized for further study.  
  
## How do I use this?
In order to identify the functional annotation of a single gene, begin with a query. Entering a single gene in the search box produces a series of tables and plots that identifies a functional map of the processes that gene *might* be involved in. In some cases, querying a gene with known functions will identify genes with well-established connections to the query gene; in other cases, new genes and new biological process might be identified, suggesting there is more to discover for well-known pathways. Querying unknown genes is especially powerful, as the associated genes and pathways provide a starting point for an otherwise difficult problem to prioritize experimentally.  
  
```{r p53_code, message=FALSE, warning=FALSE, include=FALSE}
#summary
fav_gene_summary <- gene_summary %>% 
  filter(approved_symbol == fav_gene)

#dep tables
dep_top <- make_top_table(master_top_table, fav_gene)
dep_bottom <- make_bottom_table(master_bottom_table, fav_gene)

#pathway enrichment analyses
flat_top_complete <- make_enrichment_top(master_positive, fav_gene)
flat_bottom_complete <- make_enrichment_bottom(master_negative, fav_gene)
```
#### 1. Query YFG (your favorite gene)
As an example, we will query the protein P53. Typing in the given protein name "P53" produces an error, because the official gene symbol needs to be entered. Querying TP53 (official gene symbol) generates a short summary of the gene, its name and list of aliases, and an Entrez gene summary paragraph when available.  
  
#### Summary  
**Gene**: `r fav_gene_summary$approved_symbol`  
**Name**: `r fav_gene_summary$approved_name`  
**aka**: `r fav_gene_summary$aka`  
**Entrez ID**: `r as.integer(fav_gene_summary$ncbi_gene_id)`
  
`r fav_gene_summary$entrez_summary`  
  
#### 2. Dependencies
The first plot shows the distribution of dependency scores across `r length(achilles$X1)` cell lines ranked from lowest (strongest dependencies) to highest (no dependency or inverse). Each of the `r length(achilles$X1)` cell lines is represented by a single point on the plot. Generally, values below -1 indicate the gene of interest (`r fav_gene` in this example) is essential in that cell line; values between -1 and 0, mean that cells lose fitness, but the gene is not essential; values hovering around zero indicate that ablation of `r fav_gene` has little effect on cell growth; values above 1, indicate that knocking-out the gene leads to a fitness advantage. In the case of `r fav_gene`, several cells have a fitness advantage in its absence, consistent with its role as a tumor suppressor.  
  
The second plot is a histogram of dependency scores, showing the distribution of scores for `r fav_gene`. While the majority of cells have little change in cellular growth when `r fav_gene` is absent (the histogram is centered around zero), some cells require `r fav_gene` for growth (cells scoring below -1), whereas in other cells `r fav_gene` functions as a tumor suppressor (cells with a score above 1).  
  
  
```{r dep_plots, echo=FALSE, message=FALSE, warning=FALSE}
dep_plot1 <- make_celldeps(achilles, expression_join, fav_gene, mean_virtual_achilles)
dep_plot2 <- make_cellbins(achilles, expression_join, fav_gene)
dep_plot2 + dep_plot1
ggsave(here::here("results", "fig4a_dep_plot1.png"), plot = dep_plot1, dpi = 300)
ggsave(here::here("results", "fig4b_dep_plot2.png"), plot = dep_plot2, dpi = 300)
```
  
  
To identify the cells at the tails of the plots, a dependency table will show the ranked cells by dependency score, with cell lineage information appended. In some cases, specific cell types or lineages will show consistent patterns of dependency on a gene.  
  
  
#### Cells with strong `r fav_gene` genetic dependencies:
```{r target_achilles_top, echo=FALSE, message=FALSE, warning=FALSE}
achilles_raw_long %>%  
  filter(gene == fav_gene) %>% 
  left_join(expression_join, by = "X1") %>% 
  select(cell_line, lineage, dep_score) %>% 
  top_n(dep_score, n = 10) %>% 
  mutate(dep_score = round(dep_score, 2)) %>% 
  knitr::kable(col.names = c("Cell Line", "Lineage", "Dependency Score")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```
  
  
  
#### Cells with low or inverse `r fav_gene` genetic dependencies
```{r target_achilles_bottom, echo=FALSE, message=FALSE, warning=FALSE}
achilles_raw_long %>%  
  filter(gene == fav_gene) %>% 
  left_join(expression_join, by = "X1") %>% 
  select(cell_line, lineage, dep_score) %>% 
  top_n(dep_score, n = -10) %>% 
  arrange(dep_score) %>% 
  mutate(dep_score = round(dep_score, 2)) %>% 
  knitr::kable(col.names = c("Cell Line", "Lineage", "Dependency Score")) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```
  
  
  
Understanding the shape of the curve and distribution of the raw data underlying the patterns is important for interpreting the results.    
  
  
#### 3. Similar Patterns
Positive correlations of dependency scores are ranked for each gene. Recall that these genes show similar patterns of dependencies in the same cell lines. More simply, the cells that care about `r fav_gene` deletion also care about deletion of these genes, implying a functional relationship. In the Dependency Score Example heatmap schematic above, `r fav_gene` is gene X, and genes with similar patterns would be genes A, B, and C. The `r paste0(length(dep_top$Gene))` genes that show a  similar genetic dependencies as `r fav_gene` and are above `r sd_threshold` standard deviations away from the resampled mean are displayed.  
  
One data source that can help prioritize genes for study is to compare gene-gene associations to known gene-gene relationships based on previous literature. By counting the number of times a p53 was published with each of these top ranked genes, the Relaxin family peptide receptor 2 (RXFP2) emerges as an unexpected connection, based on current literature.  
  
```{r dep_top, echo=FALSE, message=FALSE, warning=FALSE}
head(dep_top, 10) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = FALSE)
```
  
  
These `r paste0(length(dep_top$Gene))` genes were queried for gene set enrichment, and the gene sets and pathways with the strongest statistical significance are shown. Simply stated, these are the pathways that best represent the list of genes that share similar genetic dependencies, and suggest that the query gene is part of these pathways.  
  
  
```{r flat_top_complete, echo=FALSE, message=FALSE, warning=FALSE}
flat_top_complete %>% 
  slice(1:10) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = FALSE)
```
  
  
#### 4. Dissimilar
Like the analysis for genes that share similar patterns, this analysis can be used to find genes that share distinctly dissimilar patterns; that is, genes that have an inverse correlation of dependences. Simply stated, the cells that care about `r fav_gene` deletion *do not* care about deletion of these genes, implying an inverse or opposing relationship. In the Dependency Score Example heatmap schematic above, `r fav_gene` is gene X, and genes with dissimilar patterns would be genes D, E, and F. The `r paste0(length(dep_bottom$Gene))` genes that show inverse genetic dependencies to `r fav_gene` and are below `r sd_threshold` standard deviations away from the resampled mean are:   
  
  
```{r dep_bottom, echo=FALSE, message=FALSE, warning=FALSE}
head(dep_bottom, 10) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"))
```
  
  
These `r paste0(length(dep_bottom$Gene))` genes were also queried for gene set enrichment, and the gene sets and pathways with the strongest statistical significance are shown. Simply stated, these are the pathways that best represent the list of genes that have inverse genetic dependencies. 
  
  
```{r flat_bottom_complete, echo=FALSE, message=FALSE, warning=FALSE}
flat_bottom_complete %>% 
  slice(1:10) %>% 
  knitr::kable() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = FALSE)
```
  
  
How to interpret these genes and pathways is more variable than the positively correlated genes and pathways. In some cases, a negative regulator of a gene has a negative correlation with that gene, such as in this example with `r fav_gene` and its negative regulator `r dep_bottom$Gene[1]`. In other cases, opposing *pathways* are shown, contrasting the `r fav_gene` enriched pathway term "`r flat_top_complete[11,2]`" with the dissimilar enriched pathway term "`r flat_bottom_complete[11,2]`", revealing two opposing biological pathways.    
  
  
#### 5. Graph
Identifying genes that share similar patterns of dependency to a queried unknown gene generates strong hypotheses about new functional annotations and maps to new pathways. However, the strength of the hypothesis cannot be fully inferred from single gene list. If a new gene is associated with the queried gene, then you might infer a functional relationship. However, if you inspect the top 10 genes with the queried gene, then inspect the top 10 genes of each of those, building a functional network graph of the top related genes might reveal a stronger association of the new gene with your queried gene *and* its top ranked genes.  
  
```{r graph, echo=FALSE, message=FALSE, warning=FALSE}
# from fun_tables.R and fun_graphs.R
# p53graph <- make_graph_report(gene_symbol = fav_gene)

# p53graph

make_graph(master_top_table, master_bottom_table, fav_gene, 
           displayHeight = '80vh', displayWidth = '100%', tooltipLink = TRUE) %>% 
  visLegend(position = "right", width = .25, zoom = F) %>% 
  visSave(here::here("tempNetwork.html"))
# display the network
  htmltools::includeHTML(here::here("tempNetwork.html"))

# ggsave(here::here("results", "fig4c_p53graph.png"), plot = p53graph, dpi = 300)
```

```{r remove_graph, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
# Remove the temp network
file.remove(here::here("tempNetwork.html"))
```
  
The image here is a representative static, albeit busy, network graph; data presented on datadrivenhypothesis.org is an interactive image that can be zoomed, dragged, and manipulated for data exploration.

## Broader Implications
Like the proverbial man [searching for his lost keys under the lamp post](https://www.matthewhirschey.com/articles/exploratory-mind) because the light shines there, searching for biological truths often occurs under 'lamp posts' because that's where scientists can see. But what if your keys are not under the light? Or your gene is totally unknown? What do you do?  
  
The scientific method has guided scientific minds for hundreds of years, starting with a question, followed by a hypothesis, and then an experimental path to test the prediction. While hypotheses are the bedrock of science, the volume, complexity, and sophistication of modern science necessitate new methods to generate hypotheses.  
  
New tools in Data Science -- a combination of computer programming, math & statistics, and topical expertise -- combined with the rapid adoption of open science and data sharing allow scientists to access publicly available datasets and interrogate these data *before* performing any experiments.  
  
Imagine having strong data to support your new hypothesis *before* testing it. This is data-driven hypothesis.
  
## Where do I get more information? 
```{r include=FALSE}
#add publication data here first
```
Code is available on the [Hirschey Lab github account](https://github.com/matthewhirschey/ddh), including links to download the raw data, and run the analyses in R from scratch. Furthermore, the Broad Institute has a lot of information on their [website](http://www.broadinstitute.org) and the website dedicated to the [Dependency Map project](http://www.depmap.org) about how the raw data were generated, and provide a list of references.  
  
#### Code Availability  
```{r include=FALSE}
#how long
time_end_data <- Sys.time()
```
[Generate Data](https://github.com/matthewhirschey/ddh/blob/master/code/generate_depmap_data.R)  
[Statistical Analyses](https://github.com/matthewhirschey/ddh/blob/master/code/generate_depmap_stats.R)  
[Table Generator](https://github.com/matthewhirschey/ddh/blob/master/code/generate_depmap_tables.R)  
[Pathway Generator](https://github.com/matthewhirschey/ddh/blob/master/code/generate_depmap_pathways.R)  
  
`r lubridate::stamp("Methods updated December 31, 1979")(now()) `
