---
output: pdf_document
---

```{r template_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#libraries are already loaded
```

## Report
Author: Matthew Hirschey  
Printed on `r Sys.Date()`  
  
## `r str_to_title(str_replace_all(lib, "_", " "))`
**Gene Coverage:** `r as.integer(db_all[match(lib, db_all$libraryName), 1])`  
**Genes per Term:** `r as.integer(db_all[match(lib, db_all$libraryName), 2])`  
**More information:** `r db_all[match(lib, db_all$libraryName), 4]`  
  
## Positive Correlated Gene Sets from `r lib` 
Focusing on `r fav_gene`, `r length(dep_top$gene)` genes show a strong postive correlation for similar genetic dependencies  
  
```{r template_top, echo=FALSE}
if(length(list_top$Term) > 0) {
  pander(head(list_top, n = 10), justify = c("left", "right", "left"))
} else {
  pander("None")
}
```

## Negative Correlated Gene Sets from `r lib` 
Focusing on `r fav_gene`, `r length(dep_bottom$gene)` genes show a strong negative correlation for opposing genetic dependencies  
  
```{r template_bottom, echo=FALSE}
if(length(list_bottom$Term) > 0) {
  pander(head(list_bottom, n = 10), justify = c("left", "right", "left"))
} else {
  pander("None")
}
```


