---
title: "Dep Map correlation analysis"
output:
  pdf_document:
    toc: TRUE
params:
  release: "19Q3"
---
Overall goal of this project is to generate a correlation matrix from Broad Dependency Map data, to look at 'dependency signatures'. This will allow us to infer a functional relationship between genes, based on similar changes in cellular fitness. Chose to get data from Broad, as they have previous made available dependency correlation data on www.depmap.org. Will use Sanger to try to validate some of the findings; unfortunatley, Sanger lists only binary scores.

##Load libraries
```{r load_block, echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(here)
library(readxl)
library(janitor)
library(feather)
library(corrr)
library(purrr)
library(viridis)
library(enrichR)
library(glue)
library(moderndive)
library(rmarkdown)
library(gridExtra)
library(knitr)
library(pander)
library(vroom)
library(beepr) #long analysis; get some coffee, and comeback when ready

#clear environment
#rm(list=ls()) 
```
##define functions
```{r}
```

##import
Import Broad Insititute and Sanger Depedencay data; added Broad cell line expression data (Aug 21, 2019).
```{r import}
##BROAD
achilles <- read_feather(here::here("data", paste0(params$release, "_achilles.feather")))
achilles_long <- achilles %>% gather("gene", "dep_score", -X1)
achilles_cor <- read_feather(here::here("data", paste0(params$release, "_achilles_cor.feather"))) 
class(achilles_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df
achilles_cor_long <- read_feather(here::here("data", paste0(params$release, "_achilles_cor_long.feather")))

#achilles_cor_small <- achilles_cor %>% slice(1:1000) %>% select(1:1000) #to optimize code
#achilles_long_small <- achilles_long %>% sample_n(1000)


##SANGER
#data from: https://score.depmap.sanger.ac.uk/downloads
#sanger <- read_feather(here("raw", "sanger.feather"))
#sanger_small <- sanger %>% slice(1:100) %>% select(1:100) #to optimize code
#I downloaded the "Fitness/Non-fitness Binary Matrix", but this binary information was less rich than Dep Map. Perhaps try "Copy Number Bias Corrected Fold Change Values"? 

#EXPRESSION(BROAD)
expression <- read_feather(here::here("data", paste0(params$release, "_expression.feather")))
expression_long <- expression %>% gather("gene", "gene_expression", -X1) %>% arrange(desc(gene_expression))
expression_cor <- read_feather(here::here("data", paste0(params$release, "_expression_cor.feather"))) 
class(expression_cor) <- c("cor_df", "tbl_df", "tbl", "data.frame") #define class so functions (eg focus) can work on reloaded df
expression_cor_long <- read_feather(here::here("data", paste0(params$release, "_expression_cor_long.feather")))

expression_id <- read_feather(here::here("data", paste0(params$release, "_expression_id.feather"))) 

expression_join <- expression_id %>% 
  rename(X1 = dep_map_id) %>% 
  select(X1, stripped_cell_line_name, lineage)

#expression_cor_small <- expression_cor %>% 
#  select(1:10) %>%  #to optimize code
#  sample_n(1000)
```

# Global Analysis
##EDA
```{r eda}
achilles_cor_max <- achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) 

#sanger_cor_max <- sanger_cor %>% 
#  select(-rowname) %>% 
#  summarize_all(list(~max(., na.rm = TRUE))) %>% 
#  gather("gene", "max") %>% 
#  arrange(desc(max)) 

expression_cor_max <- expression_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>%  #I think there are some NAs in there...messes up the fun(max)
  gather("gene", "max") %>% 
  arrange(desc(max)) 

ggplot(achilles_cor_max, aes(max)) +
  geom_histogram(binwidth = 0.05)
#most genes have pretty low correlation
#some with high corr, TSC1/2, are only high with eachother
#others are clusters, NDUFB10
#some show bona fide pathways; CAD

#ggplot(sanger_cor_max, aes(max)) +
#  geom_histogram(binwidth = 0.05)

ggplot(expression_cor_max, aes(max)) +
  geom_histogram(binwidth = 0.05)

achilles_cor_min <- achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>% 
  gather("gene", "min") %>% 
  arrange(min) 

sanger_cor_min <- achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>% 
  gather("gene", "min") %>% 
  arrange(min) 

ggplot(achilles_cor_min, aes(min)) +
  geom_histogram(binwidth = 0.05)
#postive correlation 1: p53, p53bp1, check2, atm, usp28, etc.
#postiive correlation 2: mdm2, mdm4, ppm1d, ppm1g
#corr 1 and corr 2 are NEGATIVELY correlated with eachother; consistently.

#ggplot(sanger_cor_min, aes(min)) +
#  geom_histogram(binwidth = 0.05)
```

#Permutation tests
A permutation test involves permuting one or more variables in a data set before performing the test, in order to break any existing relationships and simulate the null hypothesis. In this case, we will break the relationship between gene-gene pairs and the correlation values. We can then compare the true statistic (mean correlation) to the generated distribution of null statistics (fake means). This will give a better idea of where to draw a threshold of a "signficiant correlation" for these analyses

Can I do this for one gene, sampling from the entire correlation matrix, and getting a distribution of correlated means for a single gene 1000x?

Consider: https://tidymodels.github.io/rsample/articles/Basics.html
#moderndive package
https://moderndive.com/8-sampling.html

```{r permutation}
  virtual_achilles <- achilles_cor_long %>% #achilles_cor_long already has all of the variables in a long format
  filter(!is.na(r)) %>%   
  rep_sample_n(size = 20000, reps = 1000) %>% #larger sample size, less error (but only 625 sets, and we're mimicking 1000?, but 310M combinations, so probably OK)
  group_by(replicate) %>% 
  summarize(mean = mean(r), max = max(r), min = min(r), sd = sd(r)) #how to handle + vs. - correlation?

#instead of filtering, could:
#summarize_all(list(max = max(r, na.rm = TRUE), 
#                     min = min(r, na.rm = TRUE),
#                     mean = mean(r, na.rm = TRUE),
#                     sd = sd(r, na.rm = TRUE))) 
#but threw and error

#The mean may not be the best measure of centrality due to its sensitivity to outliers. The example data used here seems asymmetric and it that case the median is a better measure of centrality.
#https://github.com/JoachimGoedhart/PlotsOfDifferences

ggplot(virtual_achilles, aes(x = sd)) +
  geom_histogram(color = "white") +
  labs(x = "SD r^2 value from virtual achilles", 
       title = "Distribution of mean r^2 values") 

mean_virtual_achilles <- mean(virtual_achilles$mean)
sd_virtual_achilles <- mean(virtual_achilles$sd)

sd_threshold <- 3

achilles_upper <- mean_virtual_achilles + sd_threshold*sd_virtual_achilles
achilles_lower <- mean_virtual_achilles - sd_threshold*sd_virtual_achilles
```

#Correlation network for top and bottom genes
```{r network}
#top
achilles_cor_top50 <- achilles_cor %>% 
  select(-rowname) %>% #select(one_of(target_vec)) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) %>% 
  top_n(50) 

pander(achilles_cor_top50)
  
achilles_cor_bottom50 <- achilles_cor %>% 
  select(-rowname) %>% #select(one_of(target_vec)) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>% 
  gather("gene", "min") %>% 
  arrange(min)  %>% 
  top_n(-50)

pander(achilles_cor_bottom50)

focus_50 <- c(achilles_cor_top50, achilles_cor_bottom50)

expression_cor_max %>% 
  top_n(50)

```

#global summary
```{r summary}
#top mean, sd
#achilles_cor %>% 
#  select(-rowname) %>% 
#  summarize_all(max) %>% 
#  gather("gene", "max") %>% 
#  arrange(desc(max)) %>%
#  summarize_at("max", list(mean, sd))
#-0.32 +/- 0.10
#therefore, 2 SD from min mean is 0.5264
#summarize(n()) = 918 genes
  
achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>%
  gather("gene", "max") %>% 
  arrange(desc(max)) %>%  
  top_n(50) %>% #filter(max > achilles_upper) %>%
  kable(format="html", digits = 3, caption = "Genes with large positively correlated gene dependency signatures", col.names = c("Gene", "r^2"))

#bottom mean, sd
#achilles_cor %>% 
#  select(-rowname) %>% 
#  summarize_all(min) %>% 
#  gather("gene", "min") %>% 
#  arrange(min)  %>% 
#  summarize_at("min", list(mean, sd))
#-0.23 +/- 0.03
#therefore, 2 SD from min mean is -0.2983
#summarize(n()) = 641 genes

achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>%  
  gather("gene", "min") %>% 
  arrange(min) %>%  
  top_n(-20) %>% #filter(min < achilles_lower) %>%
  kable(format="html", digits = 3, caption = "Genes with large negatively correlated gene dependency signatures", col.names = c("Gene", "r^2"))


```

#Focused Analysis
```{r import_targets}
#write_feather(id, path = here("raw", "id.feather"))
id <- read_feather(here("data", "id.feather"))

proteins <- id %>% 
  select(gene, protein_name)

SIRT <- c("SIRT1", "SIRT2", "SIRT3", "SIRT4", "SIRT5", "SIRT6", "SIRT7")
target_vec <- c("CPT1A", "CPT1B", "CPT1C", "CPT2", "SLC25A20")
hdac <- str_subset(names(achilles_cor), "HDAC")

#guest projects
trub <- c("FASN", "HMGCR", "HMGCS1")
transferase <- id %>% filter(str_detect(protein_name, "transferase")) %>% pull(gene)
acyltransferase <- id %>% filter(str_detect(protein_name, "acyltransferase")) %>% pull(gene)
deacylase <- id %>% filter(str_detect(protein_name, "deacetylase") | str_detect(protein_name, "deacylase")) %>% pull(gene)
abraham <- c("KCNK3", "KCNK9", "KCNK2")
angiotensin <- id %>% filter(str_detect(protein_name, "angiotensin")) %>% pull(gene)
myc <- id %>% filter(str_detect(gene, "MYC")) %>% pull(gene)
koves <- c("MLYCD")
klhl <- id %>% filter(str_detect(gene, "KLHL")) %>% pull(gene)
boyce <- c("GALE", "OGA", "OGT", "KEAP1", "GAN", "ENC1", klhl)
keap <- c("GAN", "ENC1")
bcaa <- c("ABAT", "ACAA1", "ACAA2", "ACAD8", "ACADM", "ACADS", "ACADSB", "ALDH1B1", "ALDH2", "ALDH3A2", "ALDH6A1", "ALDH7A1", "ALDH9A1", "AOX1", "AUH", "BCAT1", "BCAT2", "BCKDHA", "BCKDHB", "BCKDK", "DBT", "DLD", "ECHS1", "EHHADH", "HADH", "HADHA", "HADHB", "HIBADH", "HIBCH", "HMGCL", "HMGCL", "HMGCS2", "HSD17B10", "IVD", "MCCC1", "MCCC2", "MCEE", "MUT", "PCCA", "PCCB")
bcaa_trim <- c("MUT", "PCCA", "PCCB")
creatine <- c("GATM", "GAMT", "SLC6A8", "CKB", "CKM", "CKMT1B", "CKMT2")
glycolysis <- c("ALDOA", "BPGM", "ENO1", "ENO2", "GAPDH", "GPI", "HK1", "HK2", "HKDC1", "PFKL", "PFKM", "PGAM1", "PGAM2", "PGAM4", "PGK1", "PKLR", "PKM", "TPI1")
cardiolipin <- c("CKMT1A", "CKMT1B", "CYCS", "NME4", "TAZ", "TMEM256-PLSCR3", "CRLS1", "PGS1", "PTPMT1", "TAMM41")
cholesterol <- c("ACAT2", "CYP51A1", "DHCR24", "DHCR7", "EBP", "FDFT1", "FDPS", "GGPS1", "HMGCR", "HMGCS1", "HSD17B7", "IDI1", "LBR", "LIPA", "LSS", "MSMO1", "MVD", "MVK", "NSDHL", "PMVK", "SC5D", "SOAT1", "SQLE", "TM7SF2")
tca <- c("ACLY", "ACO1", "ACO2", "CS", "DLAT", "DLD", "DLST", "FH", "IDH1", "IDH2", "IDH3A", "IDH3B", "IDH3G", "MDH1", "MDH2", "MPC1", "OGDH", "OGDHL", "PC", "PCK1", "PCK2", "PDHA1", "PDHA2", "PDHB", "SDHA", "SDHB", "SDHC", "SDHD", "SUCLA2", "SUCLG1", "SUCLG2")
purine <- c("ADSL", "ADSS1", "ADSS2", "AK1", "AK2", "AK3", "AK4", "AK5", "ATIC", "GART", "GMPS", "IMPDH1", "IMPDH2", "PAICS", "PFAS", "PPAT", "PRPS1", "PRPS1L1", "PRPS2", "TAF9")
pyrimidine <- c("CAD", "CMPK1", "CMPK2", "CTPS1", "CTPS2", "DHODH", "NME1", "NME2", "NME3", "NME4", "NME6", "UMPS")
ampk <- c("CAB39", "CALM1", "CALM2", "CALM3", "CAMKK1", "CAMKK2", "PPM1E", "PPM1F", "PPP1CA", "PPP1CB", "PPP1R3D", "PPP2CA", "PRKAA1", "PRKAA2", "PRKAB1", "PRKAB2", "PRKAG1", "PRKAG2", "PRKAG3", "STK11", "STRADA")
ampk_trim <- c("PRKAA2", "PRKAB1", "PRKAB2", "PRKAG1", "PRKAG2", "PRKAG3", "STK11", "STRADA")
fao <- c("ABCD1", "ABCD2", "ACAA2", "ACAD9", "ACADL", "ACADM", "ACADS", "ACADVL", "ACAT1", "ACSL1", "ACSL3", "ACSL4", "ACSL5", "ACSL6", "ACSM1", "ACSM2A", "ACSM2B", "ACSM3", "ACSM4", "ACSS1", "ACSS2", "ACSS3", "CPT1A", "CPT1B", "CPT2", "CRAT", "CROT", "ECHS1", "ETFA", "ETFB", "ETFDH", "HADH", "HADHA", "HADHB", "HSD17B10", "PEX11G", "PEX13", "PEX14", "SLC25A20")
glycogen <- c("GBE1", "GCK", "GSK3A", "GSK3B", "GYG1", "GYS1", "GYS2", "PGM1", "PGM2", "PGM3", "PPP1CA", "PPP1CB", "PPP1CC", "PPP1R3A", "PPP1R3B", "PPP1R3C", "PPP1R3D", "PPP1R3E", "UGP2", "AGL", "CALM1", "CALM2", "CALM3", "G6PC", "PHKA1", "PHKA2", "PHKB", "PHKG1", "PHKG2", "PYGB", "PYGL")
fas <- c("ACACA", "ACACB", "ACLY", "FASN", "MCAT", "OLAH", "OXSM", "ACAA2", "ECHS1", "HADHA", "HADHB", "HSD17B10", "MECR", "PPT1")
urea <- c("ARG1", "ARG2", "ASL", "ASS1", "CPS1", "GLS2", "GLUD1", "GOT2", "GPT", "OTC", "SLC1A4", "SLC1A5", "SLC25A12", "SLC25A15")
ashley <- c("NMNAT1", "NMNAT2", "NMNAT3", "PNP")
```

#EDA to get gene lists for focused targets
```{r pull_genes}
gene <- c("SIRT4")

achilles_cor %>% 
  focus(gene) %>% 
  arrange(desc(.[[2]])) %>% #use column index
  filter(rowname == "SIRT4")

achilles_cor %>% 
  focus(gene) %>% 
  arrange(desc(.[[2]])) %>% #use column index
  top_n(50)

achilles_cor %>% 
  focus(gene) %>% 
  arrange(.[[2]]) %>% #use column index
  top_n(-50)
#this is to get neg correlators

#sanger_cor %>% 
#  focus(gene) %>% 
#  arrange(desc(.[[2]])) %>% #use column index
#  top_n(50)

#sanger_cor %>% 
#  focus(gene) %>% 
#  arrange(.[[2]]) %>% #use column index
#  top_n(-50)
#this is to get neg correlators

#expression
expression_cor %>% 
  focus(gene) %>% 
  arrange(desc(.[[2]])) %>% #use column index
  top_n(50)

expression_cor %>% 
  focus(gene) %>% 
  arrange(.[[2]]) %>% #use column index
  top_n(-50)
```

#Strongest Correlations from Targeted List
```{r strongest, warning=FALSE}
#top
achilles_cor_top <- achilles_cor %>% 
  select(one_of(SIRT)) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) %>% 
  top_n(50) 

pander(achilles_cor_top)
  
achilles_cor_bottom <- achilles_cor %>% 
  select(one_of(SIRT)) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>% 
  gather("gene", "min") %>% 
  arrange(min)  %>% 
  top_n(-50)

pander(achilles_cor_bottom)

```
#pull tops and bottoms
```{r eval=FALSE, warning=FALSE, include=FALSE}
list <- achilles_cor %>% 
  select(-1) %>% #select(one_of(SIRT)) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) %>% 
  mutate_if(is.numeric, round, 3)

write_csv(list, path = "output/full_top_list.csv")

list2 <- achilles_cor %>% 
  select(-1) %>% #select(one_of(SIRT)) %>% 
  summarize_all(list(~min(., na.rm = TRUE))) %>% 
  gather("gene", "min") %>%  
  arrange(min) %>% 
  mutate_if(is.numeric, round, 3)

write_csv(list2, path = "output/full_bottom_list.csv")
```

```{r alec_diff, eval=FALSE}
fav_gene <- c("HMGCR") #or HMGCS1

 dep_top <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(desc(.[[2]])) %>% #use column index
      filter(.[[2]] > achilles_upper) %>% #formerly top_n(20), but changed to mean +/- 3sd
      rename(gene = rowname) %>% 
      left_join(proteins, by = "gene") %>% 
      select(gene, protein_name, fav_gene) %>% 
      rename(protein = protein_name, r2 = fav_gene)
    
    dep_bottom <- achilles_cor %>% 
      focus(fav_gene) %>% 
      arrange(.[[2]]) %>% #use column index
      filter(.[[2]] < achilles_lower) %>% #formerly top_n(20), but changed to mean +/- 3sd
      rename(gene = rowname) %>% 
      left_join(proteins, by = "gene") %>% 
      select("gene", "protein_name", fav_gene) %>% 
      rename(protein = protein_name, r2 = fav_gene)
    #this is to get neg correlators

#make HMGCS1 vecs        
dep_top_hmgcs1 <- dep_top %>% pull(gene) #or HMGCS1/R
dep_bottom_hmgcs1 <- dep_bottom %>% pull(gene) #or HMGCS1/R

#filter
dep_top_diff <- dep_top %>%
  filter(gene %in% dep_top_hmgcs1 == FALSE)

dep_bottom_diff <- dep_bottom %>%
  filter(gene %in% dep_bottom_hmgcs1 == FALSE)

#enrichr
dep_top <- dep_top_diff
dep_bottom <- dep_bottom_diff
```

#integrate
```{r eval=FALSE, include=FALSE}
expression_cor_max <- expression_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>%  #I think there are some NAs in there...messes up the fun(max)
  gather("gene", "max") %>% 
  arrange(desc(max)) 

achilles_cor_max <- achilles_cor %>% 
  select(-rowname) %>% 
  summarize_all(list(~max(., na.rm = TRUE))) %>% 
  gather("gene", "max") %>% 
  arrange(desc(max)) 

joined_max_cor <- expression_cor_max %>% 
  rename(expression_max = max) %>% 
  left_join(achilles_cor_max, by = "gene") %>% 
  rename(achilles_max = max) %>% 
  mutate(rms = sqrt(((expression_max^2)+(achilles_max^2))/2))

ggplot(joined_max_cor) +
  geom_point(aes(x = expression_max, y = achilles_max, color = rms)) +
  scale_color_viridis(direction = -1) +
  theme_minimal()

ggplot(joined_max_cor, aes(x = expression_max, y = achilles_max)) +
  geom_hex(bins = 20, na.rm = TRUE, size = 0.2, color = "black") +
  scale_fill_viridis_c(option = "C") +
  coord_fixed()
  
achilles_squash <- achilles_long %>%
  unite(X1, gene, col = "match", sep = "_", remove = TRUE)

joined <- expression_long %>% 
  unite(X1, gene, col = "match", sep = "_", remove = TRUE) %>% 
  left_join(achilles_squash, by = "match") %>% 
  separate(match, into = c("X1", "gene"), sep = "_")

#plot pairs of raw cellline expression-depmap scores (expression is high and dep is high?)  
joined %>% 
  #sample_n(1000) %>% 
  ggplot(aes(x = dep_score, y = gene_expression)) + #data = joined, 
  geom_point(alpha = 0.01) +
  #stat_density_2d(geom = "raster", aes(fill = stat(density)), contour = FALSE)
  geom_vline(xintercept = 0, color = "white")

ggsave(here::here("results", "expression_v_dep.png"), plot = last_plot(), dpi = 600)

log_breaks <- c(0, 1, 10, 100, 1000, 10000, 100000, 1000000)
joined %>% 
  #sample_n(10000) %>% 
  ggplot() +
  geom_hex(aes(x = dep_score, y = gene_expression), 
           bins = 90, na.rm = TRUE, size = 0.2, color = "black") +
  scale_fill_viridis_c(option = "C", name = "Count", trans = "log10", breaks = log_breaks, labels = log_breaks) +
  labs(x = "Dependency Score", y = "Gene Expression") +  
  #coord_fixed() +
  theme_minimal()

#solution to log transform scale here:
#https://stackoverflow.com/questions/8069837/is-there-a-built-in-way-to-do-a-logarithmic-color-scale-in-ggplot2
#https://unconj.ca/blog/custom-hexbin-functions-with-ggplot.html

ggsave(here::here("results", "expression_v_dep2.png"), plot = last_plot(), dpi = 600)
#prediction was v-shaped curve, with stronger dependencies as you move up in gene expression. 
#expression has nothing to do with dep...?

#if a gene is correlated with another gene (via dependency OR via expression), capture that.
#plot correlation (abs?) of gene-gene cor long

#difference between is high v. low
#value (avg) is high v. low
```

#differential analysis
```{r}
source(here::here("code", "current_release.R"))

#import raw data
load(file=here::here("data", paste0(release, "_achilles.RData")))
#import presaved corr data
load(file=here::here("data", paste0(release, "_achilles_cor.RData")))
#make long
achilles_cor_long <- achilles_cor %>% 
  stretch()

#make cor matrix from original data
achilles_cor_original <- achilles %>% #not clean
  select(-X1) %>% 
  correlate()

achilles_cor_original_long <- achilles_cor_original %>% 
  stretch()

joined <- achilles_cor_original_long %>% 
  left_join(achilles_cor_long, by = c("x" = "x", "y" = "y"))

```

#look at concordance between achilles gene names and official gene names
```{r eval=FALSE, include=FALSE}
source(here::here("code", "current_release.R"))

load(file=here::here("data", paste0(release, "_achilles.RData")))
load(here::here("data", "gene_summary.RData"))

names <- names(achilles)
head(names)
names <- names[-1]
head(names)

head(gene_summary)

sum(!(names(achilles) %in% gene_summary$approved_symbol)) #first column name is X1

wrong_name <- names(achilles)[(!(names(achilles) %in% gene_summary$approved_symbol)) == TRUE]
wrong_name <- wrong_name[-1]
wrong_name

fix_names <- function(wrong_name) {
var <- str_which(gene_summary$aka, paste0("(?<![:alnum:])", wrong_name, "(?![:alnum:]|\\-)")) #finds index
df <- gene_summary[var,]
right_name <- df$approved_symbol
if (length(var) == 1) {
  return(right_name)
} else {
  return(wrong_name)
}
#fixes 251, leaves 11
}

wrong_name_df <- as_tibble(wrong_name)
right_name_df <- wrong_name_df %>% 
  mutate(right_name = map(value, fix_names), 
         match = if_else(value == right_name, TRUE, FALSE)) 
  
sum(right_name_df$match)


str_which(gene_summary$aka, "(?<![:alnum:])AAED1(?![:alnum:]|\\-)")
gene_summary[25614,]

#fix_names
for (name in names(achilles_skinny)) {
  if (name %in% gene_summary$approved_symbol == FALSE){
  names(achilles_skinny)[names(achilles_skinny) == name] <- fix_names(name)
} else {
  name
}
}
achilles_skinny <- achilles %>% select(1:16)
name <- "AAED1"

#it works
for (name in names(achilles)) {
  if (name %in% gene_summary$approved_symbol == FALSE){
  names(achilles)[names(achilles) == name] <- fix_names(name)
} else {
  name
}
}
sum(!(names(achilles) %in% gene_summary$approved_symbol)) #first column name is X1


#use code to display more info if unapproved/aka gene entered
paste0("(?<![:alnum:])", gene_symbol, "(?![:alnum:]|\\-)")

str_detect(gene_summary$aka, paste0("(?<![:alnum:])", PPAR, "(?![:alnum:]|\\-)"))

str_c(pull(gene_summary[str_which(gene_summary$aka, "PPAR"), 2]), collapse = ", ")

```


#print Session information for provenance and reproducibility
```{r}
utils:::print.sessionInfo(sessionInfo()[-8]) 
#You can remove an item from sessionInfo(), which is a list with a class attribute, by printing the resulting object omitting one of the list items (omitted list of packages installed, but not loaded)
```

#beep
```{r}
beep(sound = 8) #because mario is awesome
```



