---
title: "Pubtator Notebook"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(tidytext)
library(here)
library(tictoc)
library(widyr)
#rm(list = ls())
```

#import
```{r}
#read data from create_gene_summary.R
load(here::here("data", "gene_summary.RData"))

#raw <- read_tsv(here::here("data", "pubtator", "gene2pubtatorcentral"), col_names = c("pmid", "type", "concept_id", "mentions", "resource")) %>% 
#  select(pmid, concept_id, mentions) %>% 
#  mutate(match = concept_id %in% gene_summary$ncbi_gene_id) %>% 
#  filter(match == TRUE) %>% 
#  select(-match)

#saveRDS(raw, file = here::here("data", "pubtator", "raw.Rds"))
raw <- readRDS(here::here("data", "pubtator", "raw.Rds"))

#clean
#raw <- raw %>% separate(concept_id, into = c("concept_id"), sep = "[^[:digit:]]") #drop garbage
raw$concept_id <- as.numeric(raw$concept_id)
```

#import from url
This is code that went into update_gene_summary to add concept_id counts to the gene list
This will help provide context for gene-gene (concept-concept counts)
```{r}
pubtatorurl <- "ftp://ftp.ncbi.nlm.nih.gov/pub/lu/PubTatorCentral/gene2pubtatorcentral.gz"
gene2pubtator <- read_tsv(pubtatorurl, col_names = c("pmid", "type", "concept_id", "mentions", "resource")) %>% 
  select(pmid, concept_id, mentions) %>% 
  filter(concept_id %in% gene_summary$ncbi_gene_id == TRUE)

gene2pubtator$concept_id <- as.numeric(gene2pubtator$concept_id)

gene2pubtator_count <- gene2pubtator %>% 
  group_by(concept_id) %>% 
  count(sort = TRUE) %>% 
  right_join(gene_summary_raw, by = c("concept_id" = "ncbi_gene_id"))  %>% 
  rename("ncbi_gene_id" = "concept_id", 
         "concept_count" = "n") %>% 
  mutate(concept_count = replace_na(concept_count, 0))

gene2pubtator_count %>% 
  arrange(desc(concept_count))

gene_summary %>% 
  mutate(pubmed_count_rank = dplyr::percent_rank(pubmed_count)
```


#co-occurance of concepts
Count words co-occuring within sections
```{r}
#ie what two gene co-occur in a paper, across all papers
gene_pairs <- raw %>%
  #filter(!is.na(concept_id)) %>% 
  #sample_frac(0.1) %>% 
  pairwise_count(concept_id, pmid, sort = TRUE)

#fix names before making wider
gene_summary_match <- gene_summary %>% 
  select(approved_symbol, ncbi_gene_id)

gene_pairs <- gene_pairs %>% 
  #slice(1:10) %>% 
  left_join(gene_summary_match, by = c("item1" = "ncbi_gene_id")) %>% 
  left_join(gene_summary_match, by = c("item2" = "ncbi_gene_id")) %>% 
  transmute(gene_id = approved_symbol.x, y = approved_symbol.y, n)

#make wider, and (importantly) fill missing values with "0"
gene_pair_matrix <- gene_pairs %>% 
  pivot_wider(names_from = y, values_from = n, values_fill = list(n = 0))

saveRDS(gene_pair_matrix, file = here::here("data", "gene_pair_matrix.Rds"))
#readRDS(file = here::here("data", "gene_pair_matrix.Rds"))
```

#test matrix
```{r}
head(gene_pair_matrix)

gene_pair_matrix %>% 
  select(gene_id, SIRT4) %>% 
  arrange(desc(SIRT4))

gene_pair_matrix %>% 
  select(gene_id, C6orf136) %>% 
  arrange(desc(C6orf136))

gene_pair_matrix %>% 
  select(gene_id, SDHB) %>% 
  arrange(desc(SDHB))
```

#test gene pairs
```{r}
gene_pairs %>% 
  filter(target_gene == "TNF") %>% 
  arrange(desc(n))
```


#viz
```{r}
ggplot(gene_pairs) +
  geom_histogram(aes(x = log10(n)))
```

# rank
```{r}
#ntile() was too rough (too many low ranks in high bins); used percent_rank(): a number between 0 and 1 computed by rescaling min_rank to [0, 1]; importantly, this gives the same percent to identical n's, so all the 0's are the same, all the 1's are the same, etc.

gene_pairs <- gene_pairs %>% 
  #sample_n(1000) %>% 
  mutate(log2 = log2(n)) %>% 
  arrange(desc(log2))


gene_pairs <- gene_pairs %>% 
  #sample_n(1000) %>% 
  mutate(rank = percent_rank(log2)) %>% 
  arrange(desc(rank))

head(gene_pairs, n = 100)

gene_pairs %>% 
  filter(gene_id == "SIRT4") %>% 
  arrange(desc(n))

gene_pairs %>% 
  filter(gene_id == "C6orf136") %>% 
  arrange(desc(n))

gene_pairs %>% 
  filter(gene_id == "SDHB") %>% 
  arrange(desc(n))
```
I don't really like any of these; might just leave the raw numbers.

#count
Count number of times a concept_id (i.e. gene id) was associated with a paper; when other concepts are in the same paper, it is the basis for the gene co-occurance
```{r}
count <- raw %>% 
  #sample_n(1000) %>% 
  group_by(concept_id) %>% 
  summarize(count = n()) %>% 
  mutate(rank = signif(percent_rank(count), 2)) %>% 
  arrange(desc(count)) %>% 
  left_join(gene_summary_match, by = c("concept_id" = "ncbi_gene_id"))
  
head(count, n = 50)
tail(count, n = 50)

count %>% filter(approved_symbol == "C6orf136")
```

#bootstrap!!!
sample null hypothesis of data for mean, sd, 
then calculate for dataset
then z-transform
then list z-scores for pub concept co-occurance
```{r}

```

