---
output:
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(knitr)
library(pander)
library(patchwork)
source(here::here("code", "current_release.R"))
```

## Report for `r summary$identifier`: `r summary$name`  
### from datadrivenhypothesis.org  
`r lubridate::stamp("Report generated December 31, 1979")(now())`  
  
## Project Summary  
Data-driven hypothesis is a resource for identifying novel functions for human genes developed by the [Hirschey Lab](www.hirscheylab.org). This project leverages several high-quality publicly available datasets, including release `r release` of the [Dependency Map](www.depmap.org) from the [Broad Institute](www.broadinstitute.org).  
  
## Query Summary  
**Name**: `r summary$identifier`: `r summary$name`  
**Summary**: `r summary$summary`  
  
\newpage 
## Subcellular distribution of `r summary$identifier`  
  
```{r anatogram, echo=FALSE, message=FALSE, warning=FALSE}
cellanatogram
```
  
```{r anatogram_table, echo=FALSE, message=FALSE, warning=FALSE}
pander(cellanatogram_table, split.table = Inf)
```

\newpage 
## Range of `r summary$identifier` dependency scores across `r num` cell lines  
  
```{r plot1, echo=FALSE, message=FALSE, warning=FALSE}
p1 
```
  
  **`r plot_celldeps_title`** `r plot_celldeps_legend`
  
\newpage
## Distribution of Dependency Scores
```{r plot2, echo=FALSE, message=FALSE, warning=FALSE}
p2
```
    
  **`r plot_cellbins_title`** `r plot_cellbins_legend`

\newpage
## Cell Lineage Distribution of Dependency Scores
```{r plot3, echo=FALSE, message=FALSE, warning=FALSE}
p3
```
  **`r plot_celllin_title`** `r plot_celllin_legend`
\newpage
```{r plot4, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8.5}
p4
```
  **`r plot_cellsublin_title`** `r plot_cellsublin_legend`
\newpage
## Cell Line Dependency Information
### Cells that lose fitness when `r summary$identifier` is knocked out
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pander(target_achilles_bottom, split.table = Inf)
```
    
### Cells that don't rely on `r summary$identifier` or if greater than 1 gain fitness when `r summary$identifier` is knocked out
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pander(target_achilles_top, split.table = Inf)
```
  
  
\newpage  
## Positive Ranked Dependency from DepMap  
`r if(length(dep_top$Gene) > 0) {pander(paste("The", length(dep_top$Gene), "genes that show the highest postive correlation for similar genetic dependencies are listed here."))} else {}`  
```{r template_dep_top, echo=FALSE}
if(length(dep_top$Gene) > 0) {
  dep_top %>% 
    select(`Query`, `Gene`, `Name`, `Z Score`, `Co-publication Count`) %>% 
    pander(justify = c("left", "left", "left", "right", "right"), split.table = Inf)
} else {
  pander("Zero genes reached threshold")
}
```
  
\newpage    
## Positive Correlated Gene Sets  
  
```{r template_top_complete, echo=FALSE}
if(length(dep_top$Gene) > 0) {
  flat_top_complete %>% 
    select(`Query`, `Gene Set`, `Gene List`, Overlap) %>% 
    slice(1:50) %>% 
    pander(justify = c("left", "left", "left", "right"), split.table = Inf)
} else {
  pander("None")
}
```
  
\newpage
## Negative Ranked Dependency from DepMap  
`r if(length(dep_bottom$Gene) > 0) {pander(paste0("The ", length(dep_bottom$Gene), " ranked genes that show the most negative correlation for similar genetic dependencies are listed here."))} else {}`  
```{r template_dep_bottom, echo=FALSE}
if(length(dep_bottom$Gene) > 0) {
  dep_top %>% 
    select(`Query`, `Gene`, `Name`, `Z Score`, `Co-publication Count`) %>% 
    pander(justify = c("left", "left", "left", "right", "right"), split.table = Inf)
} else {
  pander("Zero genes reached threshold")
}
```
  
\newpage  
## Negative Correlated Gene Sets  
   
```{r template_bottom_complete, echo=FALSE}
if(length(dep_bottom$Gene) > 0) {
  flat_bottom_complete %>% 
    select(`Query`, `Gene Set`, `Gene List`, Overlap) %>% 
    slice(1:50) %>% 
    pander(justify = c("left", "left", "left", "right"), split.table = Inf)
} else {
  pander("None")
}
```

\newpage  
## Network Graph of `r summary$identifier`
   
```{r template_graph, echo=FALSE}
if(length(dep_bottom$Gene) > 0) {
  graph_report
} else {
  pander("Graph cannot be built")
}
```
    
  **`r graph_title`** `r graph_legend`
  

