---
output: pdf_document
---

```{r template_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#libraries are already loaded
```

## Report for `r fav_gene`: `r id[match(fav_gene, id$gene), 3]`  
### Confidential report  
Author: Matthew Hirschey  
Printed on `r Sys.Date()`  

## Summary  
**Gene**: `r fav_gene`  
**Protein**: `r id[match(fav_gene, id$gene), 3]`  
**Alt. Protein**: `r id[match(fav_gene, id$gene), 4]`  
**Entrez ID**: `r as.integer(fav_gene_entrez)`
  
```{r summary_block, include=FALSE}
if(length(summary$X1) > 0) {
  summary <- str_remove(summary$X1, "summary: ")
}   else {
  summary <- paste0("No summary entry in Entrez as of ", Sys.Date())
}
```
`r summary`  
  
\newpage 
## Distribution of `r fav_gene` expression in `r length(expression$X1)` cell lines  
  
```{r plot, echo=FALSE, message=FALSE, warning=FALSE}
grid.arrange(exp_plot1, exp_plot2, ncol = 2)
```
  
### Cells with high `r fav_gene` expression
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pander(target_expression_top, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
```
    
### Cells with low `r fav_gene` expression
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pander(target_expression_bottom, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
```
  
\newpage 
## Positive Ranked Expression from CCLE  
`r if(length(exp_top$gene) > 0) {pander(paste("The", length(exp_top$gene), "genes that show the highest postive correlation for cell line co-expression are listed here."))} else {}`  
```{r template_exp_top, echo=FALSE}
if(length(exp_top$gene) > 0) {
  pander(exp_top, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
} else {
  pander("Zero genes reached threshold")
}
```
  
\newpage    
## Positive Correlated Gene Sets  
  
```{r template_top_complete, echo=FALSE}
if(length(exp_top$gene) > 0) {
  pander(head(flat_top_complete, n = 50), justify = c("left", "left", "right"), split.table = Inf)
} else {
  pander("None")
}
```
  
\newpage
## Negative Ranked Expression from CCLE  
`r if(length(exp_bottom$gene) > 0) {pander(paste0("The ", length(exp_bottom$gene), " ranked genes that show the most negative correlation for gene co-expression are listed here."))} else {}`  
```{r template_dep_bottom, echo=FALSE}
if(length(exp_bottom$gene) > 0) {
  pander(exp_bottom, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
} else {
  pander("Zero genes reached threshold")
}
```
  
\newpage  
## Negative Correlated Gene Sets  
   
```{r template_bottom_complete, echo=FALSE}
if(length(exp_bottom$gene) > 0) {
  pander(head(flat_bottom_complete, n = 50), justify = c("left", "left", "right"), split.table = Inf)
} else {
  pander("None")
}
```
