---
output:
  pdf_document:
    latex_engine: xelatex
---

```{r template_setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(here)
library(pander)
#library(gridExtra)
```

## Report for `r fav_gene_summary$approved_symbol`: `r fav_gene_summary$approved_name`  
### Confidential report  
Author: Matthew Hirschey  
`r lubridate::stamp("Report generated December 31, 1979")(now())`  
  
## Summary  
**Gene**: `r fav_gene_summary$approved_symbol`  
**Name**: `r fav_gene_summary$approved_name`  
**aka**: `r fav_gene_summary$aka`  
**Entrez ID**: `r as.integer(fav_gene_summary$ncbi_gene_id)`
  
`r fav_gene_summary$entrez_summary`  
  
\newpage 
## Distribution of `r fav_gene` dependency scores across `r length(achilles$X1)` cell lines  
  
```{r plot, echo=FALSE, message=FALSE, warning=FALSE}
dep_plot2
dep_plot1
```
  
### Cells with strong `r fav_gene` genetic dependencies  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pander(target_achilles_bottom, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
```
    
### Cells with low or inverse `r fav_gene` genetic dependencies
```{r, echo=FALSE, message=FALSE, warning=FALSE}
pander(target_achilles_top, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
```
  
## Positive Ranked Dependency from DepMap  
`r if(length(dep_top$gene) > 0) {pander(paste("The", length(dep_top$gene), "genes that show the highest postive correlation for similar genetic dependencies are listed here."))} else {}`  
```{r template_dep_top, echo=FALSE}
if(length(dep_top$gene) > 0) {
  pander(dep_top, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
} else {
  pander("Zero genes reached threshold")
}
```
  
\newpage    
## Positive Correlated Gene Sets  
  
```{r template_top_complete, echo=FALSE}
if(length(dep_top$gene) > 0) {
  flat_top_complete %>% 
    select(enrichr, Term, Overlap) %>% 
    slice(1:50) %>% 
    pander(justify = c("left", "left", "right"), split.table = Inf)
} else {
  pander("None")
}
```
  
\newpage
## Negative Ranked Dependency from DepMap  
`r if(length(dep_bottom$gene) > 0) {pander(paste0("The ", length(dep_bottom$gene), " ranked genes that show the most negative correlation for similar genetic dependencies are listed here."))} else {}`  
```{r template_dep_bottom, echo=FALSE}
if(length(dep_bottom$gene) > 0) {
  pander(dep_bottom, justify = c("left", "left", "right"), split.cells = c("25%", "65%", "10%"))
} else {
  pander("Zero genes reached threshold")
}
```
  
\newpage  
## Negative Correlated Gene Sets  
   
```{r template_bottom_complete, echo=FALSE}
if(length(dep_bottom$gene) > 0) {
  flat_bottom_complete %>% 
    select(enrichr, Term, Overlap) %>% 
    slice(1:50) %>% 
    pander(justify = c("left", "left", "right"), split.table = Inf)
} else {
  pander("None")
}
```
