---
title: "mitocarta validation ideas"
output: html_notebook
---

```{r setup}
library(tidyverse)
library(mitocarta)
library(proteins)
#devtools::install_github("hirscheylab/omim")
library(omim)
library(RISmed) #Load the RISmed package to access pubmed api
library(xml2)
library(vroom)
load(here::here("data", "gene_summary.RData"))
```

```{r}
mitocarta <- mitocarta %>% 
  filter(mcarta2_list == 1) %>% 
  arrange(desc(mcarta2_score))

mitocarta_orf <- mitocarta %>% 
  filter(str_detect(symbol, "orf"))

#omim

```

#functions
This is the first attempt, using the function from the RISmed package
It works for a single instance, but then when I try to loop through, it fails. 
Some of the data are returned as lists, not as values. 
```{r pubmed}
pubmed_gene_count_ris <- function(gene) {
  gene <- as.character(gene)
  search_query <- EUtilsSummary(gene, type='esearch', db='pubmed', retmax = 25000)
  count <- as.character(QueryCount(search_query))
  Sys.sleep(1)
  return(count)
}

pubmed_gene_count_ris("PYURF")
pubmed_gene_count_ris("TP53")
genes <- mitocarta$symbol
genes_short <- genes[1:10]
genes_tmp <- c("TP53", "PYURF")

k <- NULL
l <- NULL
for (i in genes) {
  j <- pubmed_gene_count_ris(i)
  k <- c(k, j)
  l <- c(l, i)
}
#cbind (k, )
bind <- cbind(k, l)

```

#remake fun() because other one kept failing
Re-writing the function will give me better control over the output, and what to do if it fails. 
```{r}
#dev code
url <- 'https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi?db=pubmed&term=SIRT4&retmax=25000'
readLines(url)
tmp <- read_xml(url)
xml_attrs(xml_child(tmp, 1))
xml_name(tmp)
list <- as_list(tmp)
list[["eSearchResult"]][["Count"]][[1]]

#working function
pubmed_gene_count <- function(gene) {
  gene <- as.character(gene)
  #search_query <- EUtilsSummary(gene, type='esearch', db='pubmed', retmax = 25000)
  #count <- as.character(QueryCount(search_query))
  url <- paste0("https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi?db=pubmed&term=", gene, "&retmax=25000&email=matthew.hirschey@duke.edu&api_key=c02d48a6960a0c40134e33436870410fe107")
  tmp <- read_xml(url)
  list <- as_list(tmp)
  count <- list[["eSearchResult"]][["Count"]][[1]]
  #Sys.sleep(1)
  return(count)
}

#started writing fun() to align ncbi gene id with gene name, but then I realized I could just use gene_summary instead!!!
pubmed_symbol_lookup <- function(id) {
  id <- as.character(id)
  url <- paste0("https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi?db=gene&id=", id, "&rettype=docsum&retmode=text&api_key=c02d48a6960a0c40134e33436870410fe107")
  tmp <- vroom::vroom(url, col_names = FALSE) %>% 
    separate(X1, into = c("a", "b"), sep = "[:punct:]") %>% 
    select(b) %>% 
    slice(1) %>% 
    pull()
  }
```

```{r}
mitocarta_orf <- mitocarta %>% 
  filter(mcarta2_list == 1, 
         str_detect(symbol, "orf")) %>% 
  left_join(gene_summary, by = c("human_gene_id" ="ncbi_gene_id")) %>% 
  mutate(pubmed_approved = map_chr(approved_symbol, pubmed_gene_count), 
         pubmed_symbol = map_chr(symbol, pubmed_gene_count), 
         pubmed = if_else(pubmed_approved != pubmed_symbol, as.numeric(pubmed_approved) + as.numeric(pubmed_symbol), as.numeric(pubmed_approved))) %>% 
  select(symbol, approved_symbol, approved_name, mcarta2_score, mcarta2_fdr, mcarta2_evidence, pubmed, entrez_summary, aka, synonyms, protein_length, coexpression_gnf_n50_score, pgc_induction_score, human_gene_id, ensembl_gene_id, hgnc_id, omim_id_supplied_by_omim, uni_prot_id_supplied_by_uni_prot) %>% 
  arrange(desc(mcarta2_score)) #%>% 
  #filter(str_detect(approved_symbol, "orf"))

write_csv(mitocarta_orf, here::here("data", "mitocarta_orf.csv"))
```

#integrate with table data for correlations
```{r}
load(file=here::here("data", "master_bottom_table.RData"))
load(file=here::here("data", "master_top_table.RData"))

mitocarta_orf_final <- mitocarta_orf %>%
  left_join(master_top_table, by = c("approved_symbol" = "fav_gene")) %>% 
  left_join(master_bottom_table, by = c("approved_symbol" = "fav_gene"), suffix = c("_top", "_bottom"))
```

